### 2011-01-14 doc/jInfer/user-tutorial.tex

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{epsfig}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{import}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{26cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer User Tutorial}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle

Target audience: jInfer users. Anyone who needs to use jInfer to create DTD, XSD or another schema from existing XML documents, schemas (again: DTD, XSD, ...) or queries (such as XPath).\\

\textit{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\

\section{Overview}
\begin{enumerate}
	\item Get \& install NetBeans.
	\item Get \& install jInfer.
	\item Infer schema.
\end{enumerate}

\section{NetBeans}
\textit{Do this if you don’t have any NetBeans installed. If you have an older version than 6.9, jInfer will not run - please update it first.}
\begin{enumerate}
	\item Go to http://netbeans.org/
	\item Download NetBeans 6.9+, whichever bundle will do.
	\item Follow installation instructions.
\end{enumerate}

\section{jInfer}
\textit{Do this if you have the correct version of NetBeans installed. Note that jInfer features automatic update - you don’t need to follow these steps each time new version comes out.}

\begin{enumerate}
	\item Go to https://sourceforge.net/projects/jinfer/files/
	\item Download the latest jInfer binaries (jInfer-X.Y-bin.zip).
	\item Unpack the ZIP, you’ll get a few NBM files.
	\item Start NetBeans, open \textit{Tools} > \textit{Plugins} from the main menu.
	\item Switch to \textit{Downloaded} tab.
	\item Click \textit{Add Plugins...}.
	\item Select all the unpacked NBM files.
	\item Click \textit{Install} and follow the instructions.
	\item Restart NetBeans.
\end{enumerate}

\section{Basic Inference}
\textit{This section assumes you have NetBeans and jInfer installed. After the first run, a jInfer Welcome window will appear, guiding you through basically the same steps that will be covered here. You might want to follow them instead.}\\

Let’s assume that you have one or more XML files, and you want to create XSD schema from them.\\

\begin{enumerate}
	\item Create a new jInfer project.
	\begin{enumerate}
		\item \textit{File} > \textit{New Project} \textbf{or} \textit{New Project} icon from the toolbar \textbf{or} right-click the area in \textit{Projects} window > \textit{New Project}.
		\item Select \textit{jInfer} category, \textit{jInfer Project}.
		\item Click \textit{Next}.
		\item \label{project-location} Choose a name (for example MyInference) and location of your project.
		\item Click \textit{Next}.
		\item In the \textit{Schema Generator} combo box select \textit{Basic XSD Exporter}.
		\item Click \textit{Finish}.
	\end{enumerate}	
	\item Add input XML files to this project.
	\begin{enumerate}
		\item Right click the newly created project in \textit{Projects} window.
		\item Select \textit{Add files}.
		\item Select XML files you want to use for schema inference.
		\item If you now open the \textit{XML} folder in your project, you will see the added files.
	\end{enumerate}	
	\item Run inference.
	\begin{enumerate}
		\item \textit{Run Project} in the toolbar (green play icon) \textbf{or} right click the project, select \textit{Run}.
		\item Generated schema will open in a new \textit{Editor} window. It will also appear in the \textit{Output} folder.
		\item You will find the file with the schema in your project (location from step \ref{project-location}, folder \textit{output}).
	\end{enumerate}
\end{enumerate}

\section{Project files}
\textit{This section assumes you have successfully inferred at least one schema.}
\begin{itemize}
	\item To add input files to a specific folder (XML documents, schemas, queries) without relying on jInfer to guess the folder from the extension, right click the folder and select \textit{Add XYZ files}.
	\item Note that input folders are virtual - files in them still reside at their original locations. You can’t actually delete them from disk from within jInfer, just remove them from these virtual folders by selecting \textit{Delete} from their context menu.
	\item \textit{Output} folder is, on the other hand, a faithful representation of the folder \textit{output} inside your jInfer project (step \ref{project-location}). If you delete a schema here, it will be deleted from the disk too.
	\item You might want to see a diff between two input documents, or old and inferred schema - just select both files at once, right-click and select \textit{Tools} > \textit{Diff} from their context menu.
\end{itemize}

\section{Setting up jInfer}
\textit{This section assumes that you have successfully inferred at least one schema and know how to manipulate input and output files in a project.}\\

To change jInfer settings, you have to first realize whether it is a change affecting a single project at a time (for example, changing the output language from XSD to DTD) or whole jInfer at once (whether schema should open in Editor window after the inference finishes).
\begin{enumerate}
	\item First type of settings is found in so called Project Preferences. You can access them from the respective jInfer project’s context menu: right click the project and select \textit{Properties}.
	\item Second type is called jInfer Options and can be found among NetBeans’ options: open them (Windows, Linux: \textit{Tools} > \textit{Preferences...}, Mac: \textit{NetBeans} > \textit{Preferences...}) and switch to \textit{jInfer} tab.
\end{enumerate}

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}