### 2011-07-09 doc/theses/svirec/kap1.tex
another commit of thesis

\chapter{Introduction}

\section{Aim of the Thesis}

\section{Structure of the Thesis}