### 2011-07-10 doc/theses/svirec/kap2.tex
another part of thesis

\chapter{Basic Definitions}

In this chapter basic definitions necessary in the following chapters are placed. This definitions are cited from \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.

\section{XML Trees and DTDs}
\begin{define}[Tree]
Given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is an set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.
\end{define}

\noindent Let also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined over an alphabet of node labels $\Sigma$ as $T_\Sigma$.

\begin{define}[XML Tree]
{\sl XML tree} is a pair $XT=\langle T,\delta \rangle$, where:
	\begin{enumerate}
		\item $T = (r, N, E, \lambda)$ is a tree in $T_{\tau\cup\alpha\cup\{S\}}$, where $\tau$ is a tag alphabet, $\alpha$ is an attribute name alphabet and $S$ is a symbol not belonging to $\tau\cup\alpha$ (representing \texttt{\#PCDATA} content of elements)$;$
		\item given a node $n$ of $T$, $\lambda(n) \in \alpha \cup \{S\} \Leftrightarrow n \in Leaves(T);$
		\item $\delta : Leaves(T) \rightarrow Str$, where $Str$ is a string alphabet, is a function associating a (string) value to every leaf of $T$.
	\end{enumerate}\qed
\end{define}

\begin{example}\label{example1ref}
Consider the following XML document representing a collection of books, and is graphically represented as an XML tree in Fig. \ref{example1}.
\begin{verbatim}
<bib>
  <book>
    <written_by>
      <author ano="A1">
        <name>John Writer</name>
      </author>
      <author ano="A1">
        <name>Eric Seller</name>
      </author>
    </written_by>
    <title>Some title</title>
  </book>
  <book>
    <written_by>
      <author ano="A2">
        <name>Adam Publisher</name>
      </author>
    </written_by>
    <title>Some title 2</title>
  </book>
</bib>
\end{verbatim}

The nodes of an XML tree have a label denoting the tag name of the element and unique element identifier in brackets. Leaf nodes are either an attribute or the textual content of an element. The label of a textual node contains string contained inside of element and unique element identifier. The label of an attribute node contains in addition to textual node a name of the attribute.\qed
\end{example}

\begin{figure}[h]
    \centering\includegraphics[scale=\myscale]{example1-new}
	\caption{An XML Tree} \label{example1}
\end{figure}

\begin{define}[DTD]\label{dtdDef}
{\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where: i) $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}; ii) $P$ is the set of \emph{element type definitions}; iii) R is the set of \emph{attribute lists}; iv) $rt \in \tau$ is the tag of the document root element.\qed
\end{define}

\begin{define}[Path Expression]
A path expression is an expression of the form $p = ('/' | '//')s_1 \dots ('/'|'//')s_m$ where $s_1, \dots, s_{m-1} \in \tau$, and $s_m \in \tau \cup \alpha \cup \{S\}$.
\end{define}

\begin{define}[Path]
{\sl Path} p on a DTD $D = (\tau, \alpha, P, R, rt)$ is a sequence $p = s_1, \dots, s_m$ of symbols in $\tau \cup \alpha \cup \{S\}$ such that:
	\begin{enumerate}
		\item $s_1=rt$;
		\item for each $i$ in $2..m-1$, $s_i \in \tau$ and $s_i$ appears in the element type definition of $s_{i-1}$;
		\item $s_m \in \alpha \Rightarrow s_m$ appears in the attribute list of $s_{m-1}$;
		\item $s_m \in \tau \cup \{S\} \Rightarrow s_m$ appears in the element type definition of $s_{m-1}$.
	\end{enumerate}\qed
\end{define}

As we have defined a \emph{path} on a DTD $D$, let us denote as $paths(D)$ the set of paths which can be defined on a DTD $D$. Also another important notion is $p(XT)$ (or $\{[\![p]\!]\}$), which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$. The set of nodes reachable from a node $v$ following path $p$ is denoted as $\{v[\![p]\!]\}$. When there is only one node in $\{v[\![p]\!]\}$, we use $v[\![p]\!]$ to denote this node. Moreover, let denote $XT.p$ the \emph{answer} of the path $p$ applied on $XT$ , that is:
\begin{itemize}
	\item[-] if $p \in EPath(D)$, where $EPath(D)$ denotes the set of the paths whose last symbol denotes an element, then $XT.p = p(XT)$
	\item[-] if $p \in StrPath(D)$, where $StrPath(D)$ denotes the set of paths whose last symbol denotes either the textual content of an element or an
attribute, then $XT.p = \{\delta_T(x)|x \in p(XT)\}$.
\end{itemize}

\section{Integrity Constraints}

Before defining of XML integrity constraint, let us introduce some notation used in the definition. A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be safe if all variables in $C$ are \emph{range restricted}, i.e. if
\begin{itemize}
 	\item for every $[X_i]p_i[Y_i]$, either $X_i$ is a constant (node identifier of a string), or there is some $[X_j]p_j[Y_j]$ in $C$ where $X_j$ is range restricted;
    \item for every built-in term $U_i\theta_i V_i$ occurring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
 \end{itemize}
A rootless tree formula is an expression of the form $p(\Phi_1 \land \dots \land \Phi_k)$ where $\Phi_i$ is a rootless path formula (expression of the form $p[y]$ where $p$ is a path expression and $y$ is a term) or a rootless tree formula and $p$ is a path expression. A tree atom is an expression of the form $[x]T$ where $T$ is a rootless tree formula and $x$ is a term.

\begin{define}[XML Integrity Constraint]\label{integConstr}
An XML constraint is a formula of the form: $$(\forall X)[\Phi(X)\supset (\exists Y_1)\Psi_1(X,Y_1) \lor \dots \lor (\exists Y_k)(X,Y_k)]$$
where $X,Y_1,\dots, Y_k$ denote distinct sets of universally and existentially quantified variables, $\Phi(X)$ and $\Phi(X) \land \Psi_i(X, Y_i) (\forall i \in [1..k])$ are safe conjunctions of built-in and tree atoms.\qed
\end{define}

\section{Functional Dependencies}

Since different approaches uses slightly different representation of an XML document, also definition of functional dependencies differs. This is reason we introduce in this section two different definitions.

\subsection{Functional Dependency 1}

In a relational database $DB$, a correspondence between values $A$ and $B$ in the tuple of $D$, models the functional dependency denoted as $A \rightarrow B$. Since in XML there is not such a standard tuple concept, the concept of XML tree tuples is introduced, corresponding to relational databases concept of tuples.

\begin{define}[Tree Tuple]\label{treeTuple}
Given an XML tree $XT$ conforming the DTD $D$, a tree tuple $t$ of $XT$ is a maximal sub-tree of $XT$ such that, for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
\end{define}

\begin{define}[Functional Dependency]\label{fd1}
Given a DTD $D$, a functional dependency on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
\end{define}

\begin{example}
Consider the XML tree of Fig \ref{example1}. The following functional dependency express the constraint that two distinct authors of the same book cannot have the same value of attribute \texttt{ano}:$$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$\qed
\end{example}

\subsection{Functional Dependency 2}

\begin{define}[Functional Dependency]\label{fd2}
With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$. Here $P$ is a  root path (path where first element is a root element of an XML document), or $P = \epsilon$ (empty path). Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$. Given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
\end{define}