### 2011-07-16 doc/theses/svirec/attachments.tex
added implementation chapter

\chapwithtoc{Attachments}

\begin{figure}[h]
\caption{XML Schema for weights}
\label{weightschema}
\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">

<xs:complexType name="Tweight">
  <xs:sequence>
    <xs:element name="path" type="xs:string"/>
    <xs:element name="value" type="xs:decimal"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tweights">
  <xs:sequence>
    <xs:element name="weight" type="Tweight"
        minOccurs="1" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:element name="weights" type="Tweights"/>
</xs:schema>
\end{verbatim}
\end{figure}

\begin{figure}[h]
\caption{XML Schema for functional dependencies}
\label{fdschema}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">

<xs:complexType name="TleftSidePaths">
  <xs:sequence>
    <xs:element name="path" type="xs:string"
        minOccurs="1" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="TrightSidePaths">
  <xs:sequence>
    <xs:element name="path" type="xs:string"
        minOccurs="1" maxOccurs="1"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tdependency">
  <xs:sequence>
    <xs:element name="leftSidePaths" type="TleftSidePaths"/>
    <xs:element name="rightSidePaths" type="TrightSidePaths"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tdependencies">
  <xs:sequence>
    <xs:element name="dependency" type="Tdependency"
        minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:element name="dependencies" type="Tdependencies"/>
</xs:schema>
\end{verbatim}
\end{figure}