### 2011-07-16 doc/theses/svirec/kap7.tex

\chapter{Conclusion}

The aim of this thesis was a proposal and implementation of algorithm repairing XML functional dependencies violations. At first, existing solutions were analyzed and described.

\section{Future Work}