### 2011-07-27 doc/theses/svirec/kap1.tex
corrections

\chapter{Introduction}

Nowadays, one of the most largely used standards for information representation and data exchange on the Internet is the eXtensible Markup Language (XML) \cite{xml}. While XML as a markup language provides syntactic flexibility, the structure of an XML document is described with a so-called XML schema language. The two most popular XML schema languages proposed by W3C are Document Type Definition (DTD) \cite{xml} and XML Schema Definition (XSD) \cite{xsd1,xsd2,xsd3}. They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness XML integrity constraints for XML \cite{ic} have been defined. Several different classes of integrity constraints have been defined for XML, however most common semantic constraint used in relational databases is functional dependency (FD) \cite{fd}.

With a defined set of functional dependencies in an XML document, one can check if an XML document satisfies defined functional dependencies \cite{satifFD}. However, data in the real world is typically dirty and often violate (does not satisfy) the set of functional dependencies and hence are inconsistent. This observation leads to a need of an algorithm to compute a repair applied to an XML document such that functional dependencies will be satisfied  again \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.

\section{Aim of the Thesis}

The aim of the thesis is to propose an algorithm repairing XML functional dependency violations for a given XML document and a set of functional dependencies. Analyses recent approaches and discusses their advantages and disadvantages. The focus of the thesis is to incorporate the weight model and to involve the user into the process of finding and applying the repair and find out how this interaction helps. A part of this thesis is an experimental implementation of the proposed algorithm and its experimental evaluation.

\section{Structure of the Thesis}

Chapter 2 introduces basic definitions which are necessary for further chapters. Chapter 3 presents recent approaches of repairing XML functional dependency violations. In Chapter 4 is the main algorithm of this thesis proposed. Chapter 5 contains details of the experimental implementation. Experimental result are presented in Chapter 6. Finally, Chapter 7 contains a conclusion and several suggestions for future work.