### 2011-07-29 doc/theses/svirec/appendixA.tex
content of CD

\chapter{Content of CD}

The CD is a part of this thesis. It contains text of this thesis, source code of jInfer framework with \texttt{FDRepairer} extension together with documentation and built jInfer as \texttt{nbm} plugins applicable into NetBeans platform. It aslo contains experimental data and results. CD has the following structure:

\begin{itemize}
	\item \textit{content.txt} - A file with this text
    \item \textit{text} - A directory with the PDF file containing text of this thesis
    \item \textit{src} - Source codes of jInfer including \texttt{FDRepairer} extension
    \item \textit{javadoc} - A generated javadoc documentation
    \item \textit{bin} - A directory with \texttt{nbm} plugins of jInfer and FDRepairer
    \item \textit{datasets} - Experimental data and results
\end{itemize}