### 2011-08-16 doc/theses/vitasek/appendix-b-iss.tex
Thesis: splitting tex files.

\chapter{IDSetSearch}
\label{appendix-iss}

TODO talk a bit about the structure of ISS module.

TODO name of the module, its place in jInfer, interface with the framework (service providing)

TODO package name root

TODO "Attribute Stats"

TODO heuristics - interfaces, construction, improvement

TODO experiments - quality, termination, parameters, experiment sets

TODO make sure there was a place where we described the model data model:)