### 2011-08-16 doc/theses/vitasek/chapter-5-future.tex

\chapter{Future work}

TODO

\begin{itemize}
  \item extension to more documents at once
  \item more heuristics and their chaining
  \item ACO, genetic, ...
  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
  \item jInfer is opensource, thus easy to extend!
\end{itemize}