### 2011-08-16 doc/theses/vitasek/chapter-6-conclusion.tex

\chapwithtoc{Conclusion}

TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them. We discussed the approach chosen in FIDAX. Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t. weight) ID set.

However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics. By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.