### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

\section{Experimental Results}

\subsection{Grammar and model generation}

%         NB class GrammarModelTiming

The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & all official and sized test data sets \\
  Iterations        & 50 \\
  Pool size         & not applicable \\
  $\alpha$, $\beta$ & not applicable \\
  \hline
\end{tabular}
\bigskip
\end{center}

The experimental set will contain $50 * (11 + 11) = 1100$ configurations: 50 iterations for 11 test data sets plus 11 sized test data sets. There will be no CHs or IHs. We will be gathering the timing data for IG extraction and model generation in GnuPlot format.

\begin{table}
  \caption{Grammar Extraction and Model Creation Times}
  \bigskip
  \label{table-experiments-grammar-model-timing}
  \centering
  \begin{tabular}{l || l | l || l | l || l | l}
    Data set & GE & GE & MC & MC & Tot & Tot \\
     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
    \hline
    \dataset{OVA1}     & < 10 & - & < 10 & - & < 10 & - \\
    \dataset{OVA2}     & < 10 & - & < 10 & - & < 10 & - \\
    \dataset{OVA3}     & 42.94 & 19.8509 & 60.92 & 27.0848 & 103.86	& 31.6911 \\
    \dataset{XMA-c}    & 140.32 &	33.2618 &	90.24 &	45.8803 & 230.56 & 56.2633 \\
	\dataset{XMA-p}    & 7518.82 &	922.8882 &	10135.46 &	502.8997 & 17654.28	& 1353.8794 \\
	\dataset{XMD}      & 979.18 &	307.1760 &	563.04 & 341.4697 & 1542.22	& 134.6883 \\
	\dataset{MSH}      & 570.24 &	167.1119 &	225.48 &	90.6775 & 795.72 & 161.8340 \\ 
	\dataset{NTH}      & 328.36 & 118.3766 &	1074.9 &	155.5604 & 1403.26 & 137.8695 \\
	\dataset{100-100}  & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{100-200}  & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{100-1000} & 18.34 & 10.2372 & 18.84 & 1.0373 & 37.18 & 9.9338 \\
	\dataset{0-0}      & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{10-5}     & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{20-20}    & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{30-45}    & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{40-80}    & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{50-125}   & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{60-180}   & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{70-245}   & < 10 & - & < 10 & - & < 10 & - \\
	\dataset{80-320}   & < 10 & - & < 10 & - & 12.48 & 8.3574 \\
	\dataset{90-405}   & < 10 & - & < 10 & - & 15.88 & 10.3778 \\
	\dataset{100-500}  & < 10 & - & < 10 & - & 18.74 &	8.8889 \\
  \end{tabular}
\end{table}

Results are in Table \ref{table-experiments-grammar-model-timing}. We are presenting the average grammar extraction (GE) times and their standard deviation, the same for model creation (MC) and total (sum of these two, Tot) times. For many data sets the average time is less than 10 ms: this is not enough to be precise and we don't calculate the standard deviation in these cases.

We can see from the results that for most data sets their model can easily be created under around one second, only in case of the biggest set \dataset{XMA-p} (13 MB) this takes some 17 seconds. We can conclude that grammar and model creation times are not a bottleneck for now. Heuristics run times will be order of magnitude higher.

\subsubsection{GLPK interface timing}

%         NB class GlpkInterfaceTiming

A related problem is how long it takes to create input for GLPK and then parse its results. We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.

\begin{table}
  \caption{GLPK Interface Times}
  \bigskip
  \label{table-experiments-glpk-timing}
  \centering
  \begin{tabular}{l || l | l || l | l || l | l}
	Data set & IC & IC & OP & OP & Tot & Tot \\
     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
	\hline
	\dataset{OVA1} & 36.46 & 66.8517 & 49.8 & 114.0687 & 86.26 & 150.1044 \\
	\dataset{OVA2} & 39.52 & 75.8210 & 48.8 & 102.4484 & 88.32 & 154.9596 \\
	\dataset{OVA3} & 34.1 & 74.1838 & 38.62 & 89.3772 & 72.72 & 134.7295 \\
	\dataset{XMA-c} & 40.88 & 88.6632 & 33.84 & 65.8636 & 74.72 & 127.7338 \\
	\dataset{XMA-p} & 36.54 & 70.7436 & 49.24 & 101.2412 & 85.78 & 145.2092 \\
	\dataset{XMD} & 37.98 & 69.2719 & 32.88 & 70.2173 & 70.86 & 114.6692 \\
	\dataset{MSH} & 40.42 & 91.9885 & 36.52 & 72.1018 & 76.94 & 138.6198 \\
	\dataset{NTH} & 36.02 & 66.3403 & 38.06 & 88.8244 & 74.08 & 128.9974 \\
	\dataset{100-100} & 46.5 & 103.3929 & 46.92 & 89.7049 & 93.42 & 158.7267 \\
	\dataset{100-200} & 42.34 & 96.1204 & 38.22 & 90.0284 & 80.56 & 152.6534 \\
	\dataset{100-1000} & 32.92 & 64.4534 & 42.1 & 89.4546 & 75.02 & 127.8541 \\
	\dataset{0-0} & 46.8 & 123.5183 & 46.92 & 102.2601 & 93.72 & 181.5228 \\
	\dataset{10-5} & 40.06 & 75.7370 & 40.1 & 72.4851 & 80.16 & 126.7135 \\
	\dataset{20-20} & 33.72 & 70.7263 & 34.1 & 66.2781 & 67.82 & 116.3783 \\
	\dataset{30-45} & 38.26 & 71.7549 & 45.94 & 110.1284 & 84.2 & 155.7594 \\
	\dataset{40-80} & 37.06 & 67.0024 & 49.26 & 106.3185 & 86.32 & 144.9918 \\
	\dataset{50-125} & 50.44 & 101.9162 & 84.76 & 364.7350 & 135.2 & 378.7835 \\
	\dataset{60-180} & 38.38 & 89.3379 & 42.54 & 94.3742 & 80.92 & 149.6049 \\
	\dataset{70-245} & 41.5 & 93.2951 & 40.3 & 93.4858 & 81.8 & 149.6797 \\
	\dataset{80-320} & 51.92 & 121.9812 & 47.98 & 96.0904 & 99.9 & 171.4617 \\
	\dataset{90-405} & 40.5 & 91.5373 & 36.46 & 88.5099 & 76.96 & 144.2890 \\
	\dataset{100-500} & 37.82 & 85.7571 & 43.4 & 90.3257 & 81.22 & 141.9103 \\
  \end{tabular}
\end{table}

Results are in table \ref{table-experiments-glpk-timing}. For each data set there are the times of (GLPK) input creation (IC) - average and standard deviation, then the same for output parsing (OP) and total (Tot).

Interestingly enough, in most cases the times to create an input for GLPK and then to parse its output are very similar. Also, for sized test data sets it is interesting to note that even though the $|V|$ and $|E|$ counts are increasing, the times remain almost the same. This probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the file it produced, and these times are probably the most relevant.

\subsection{GLPK: native vs. Cygwin}

%         NB class TimeQuality
%         NB class TimeTillOptimum

In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results. The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution, % TODO link
or via Cygwin as one of its packages.

If we find out which of these Cygwin version is better, we will be using it exclusively knowing this should not affect any other aspect of our experiments. We might also find that there is no relevant difference, which would be and interesting finding, too.

Apart from comparing different versions, we shall see how the pure GLPK approach behaves. The first part of this experiment will be limiting the run time, thus making it an instance of \heu{Truncated branch \& bound}. In this case we will see the dependency between the run time and the quality achieved in it. In the second time we will let GLPK run until optimum is found. We shall see the dependency between input size and run time needed to achieve the optimum.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & \dataset{100-500} \\
  Iterations        & 50 \\
  Pool size         & 1 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

Our experimental set will contain 500 experimental configurations for each of these two GLPK version. Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}). There will be no improvement heuristic. The only data we gather in the GnuPlot file are the final qualities (weights). The data set used is \dataset{100-500} as the biggest one in sized test data.

\begin{algorithm}
\caption{GLPK: native vs. Cygwin set generation 1}
\label{listing-experiment-glpk-native-vs-cygwin-1}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$i = 1 \to 50$}
	\FOR{$time = 1 \to 46$ step $5$}
    \STATE $ES \gets ES \cup {CH = \heu{Glpk}(limit = time), IH = \emptyset}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{figure}
  \caption{Time vs. Quality}
  \label{image-experiment-time-vs-quality}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/time-vs-quality}
\end{figure}

Results are in Figure \ref{image-experiment-time-vs-quality}. They should be interpreted as follows: for each time limit from 1 to 46 seconds there are two boxplots next to each other, the left, dashed one is the native GLPK, the right, solid one is the Cygwin GLPK. This is reflected in the tics on the X (time) axis, meaning that the axis cannot be interpreted in the usual way.

We can see from the graph that even though for smaller times (1 and 6 seconds, respectively) the Cygwin GLPK is reaching better qualities with smaller variance, starting from 11 seconds the native GLPK is at least as good or better for every following time. The results are inconclusive though, it is necessary to wait for confirmation from the second part of this experiment.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & all sized test data sets \\
  Iterations        & 50 \\
  Pool size         & 1 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

The other way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size. This experimental set will contain 550 configurations for each version. Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum. This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}). There will again be no IH, the only data we will collect are the times of the CH run in each case.

\begin{algorithm}
\caption{GLPK: native vs. Cygwin set generation 2}
\label{listing-experiment-glpk-native-vs-cygwin-2}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$i = 1 \to 50$}
	\FOR{$file \in $ sized test data}
    \STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(no\:limit), IH = \emptyset\}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{figure}
  \caption{Time Until Optimum}
  \label{image-experiment-time-until-optimum}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/time-till-optimum}
\end{figure}

Results are in Figure \ref{image-experiment-time-until-optimum}, please take a note that the Y axis is in log scale. As with the previous case, the X axis cannot be interpreted in the usual way. For each data set there are two boxplots next to each other: the left one is the native GLPK, the right one is the Cygwin GLPK.

From these results it becomes clear that the native GLPK has in general shorter running times for each and every input data set than its Cygwin counterpart. This becomes less extreme with the increasing input size, which leads us to suspicion that the core parts of computation in both cases are equally powerful. Regardless of that, we shall be using the \textbf{native} GLPK for following experiments.

To conclude the first timing experiments we introduce a summary pie chart in Figure \ref{image-experiment-timing-summary}. This shows the typical distribution of times needed to find the optimum for the \dataset{OVA3} data set.

These experiments proved that for bigger data sets the times to reach the optimum might become too long. We shall attempt to find heuristics to reach the optimum faster in the following experiments.

% TODO fix this chart
\begin{figure}
  \caption{Timing Summary}
  \label{image-experiment-timing-summary}
  \centering
    \includegraphics[width=.6\textwidth]{images/experiments/timing-pie}
\end{figure}

\subsection{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX}}

%         NB class RandomVsFuzzyVsFidaxStart

Our investigation into various CHs will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data set    & all official test data sets \\
  Iterations        & 50 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

The experimental set will contain 1650 configurations in total: 3 different CHs * 11 official test data sets * 50 iterations. There will be no improvement heuristics. The pool size will be set to 10, even though \heu{FIDAX} cannot not profit from this. Listing for this can be found in \ref{listing-experiment-random-fuzzy-fidax}.

We will be gathering the running time of the CH itself and quality of the best solution found for GnuPlot.

\begin{algorithm}
\caption{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX} set generation}
\label{listing-experiment-random-fuzzy-fidax}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$i = 1 \to 50$}
	\FOR{$file \in $ official test data}
    \STATE $ES \gets ES \cup \{file, CH = \heu{Random}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{Fuzzy}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{figure}
  \caption{Random vs. Fuzzy vs. FIDAX - Quality}
  \label{image-experiment-random-fuzzy-fidax-quality}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-quality}
\end{figure}

\begin{figure}
  \caption{Random vs. Fuzzy vs. FIDAX - Time}
  \label{image-experiment-random-fuzzy-fidax-time}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-time}
\end{figure}

Results can be found in Figure \ref{image-experiment-random-fuzzy-fidax-quality} - qualities achieved and Figure \ref{image-experiment-random-fuzzy-fidax-time} - times spent. The Y (time) axis in the latter figure is again in log scale. For each data set there are 3 boxplots next to each other. The first, leftmost, represents \heu{Random}, second \heu{Fuzzy} and finally the third, rightmost is \heu{FIDAX}.

We can draw the following conclusions. \heu{Fuzzy} consistently finds the best solution, but it's by far the slowest of these CHs. The trivial \heu{Random} is better than \heu{FIDAX} in artificial as well as some real data.

\subsubsection{Improving \heu{FIDAX} with \heu{Hungry}}

%         NB class FidaxWithHungry

Now we shall try to answer a minor question, whether it is possible to improve \heu{FIDAX} by using \heu{Hungry} as IH. This short experiment answers that question.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data set    & all official test data sets \\
  Iterations        & 1 \\
  Pool size         & 1 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

We need a pool size of one and only a single iteration - both \heu{FIDAX} and \heu{Hungry} are deterministic. We will try all official data sets, first with empty IH, second with \heu{Hungry} as IH. We will gather the qualities in each case and see whether there is any improvement.

The experimental results are summarized in the Table \ref{table-experiments-fidax-and-hungry} and are quite surprising. As trivial a heuristic \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).

Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}. Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute".

\begin{table}
  \caption{Results of adding \heu{Hungry} after \heu{FIDAX}}
  \bigskip
  \label{table-experiments-fidax-and-hungry}
  \centering
  \begin{tabular}{l | l | l}
    Data set & Quality - \heu{FIDAX} & Quality - \heu{FIDAX} + \heu{Hungry} \\
    \hline
    \dataset{OVA1}     & 0.4411764705882353  & 0.4411764705882353   \\
    \dataset{OVA2}     & 0.16346153846153846 & 0.16346153846153846  \\
    \dataset{OVA3}     & 0.25482414123443264 & \textbf{0.2553715615163541}   \\
    \dataset{XMA-c}    & 0.7546666666666666	 & 0.7546666666666666   \\
    \dataset{XMA-p}    & 0.2019306150568969	 & 0.2019306150568969   \\
    \dataset{XMD}      & 0.09786094165493509 & 0.09786094165493509  \\
    \dataset{MSH}      & 0.5416472778036296	 & 0.5416472778036296   \\
    \dataset{NTH}      & 0.05259709474828076 & \textbf{0.057918595422124436} \\
    \dataset{100-100}  & 0.56	               & \textbf{0.6766666666666669}   \\
    \dataset{100-200}  & 0.44200000000000017 & \textbf{0.5980000000000003}   \\
    \dataset{100-1000} & 0.19952380952380955 & \textbf{0.29619047619047617}  \\
  \end{tabular}
\end{table}

% TODO for some reason \texttt{\textbf{vertex5}} looks only like \texttt{vertex5}
\begin{table}
  \caption{ID sets in \heu{FIDAX} versus \heu{FIDAX} + \heu{Hungry}}
  \bigskip
  \label{table-experiments-fidax-and-hungry-idsets}
  \centering
  \begin{tabular}{l | l}
  \heu{FIDAX} & \heu{FIDAX} + \heu{Hungry} \\
  \hline
                    & \texttt{\textbf{vertex5}}  \\
                    & \texttt{\textbf{vertex26}} \\
  \texttt{vertex30} & \texttt{vertex30} \\
  \texttt{vertex31} & \texttt{vertex31} \\
  \texttt{vertex32} & \texttt{vertex32} \\
  \texttt{vertex34} & \texttt{vertex34} \\
  \texttt{vertex35} & \texttt{vertex35} \\
  \texttt{vertex36} & \texttt{vertex36} \\
  \texttt{vertex37} & \texttt{vertex37} \\
  \texttt{vertex39} & \texttt{vertex39} \\
                    & \texttt{\textbf{vertex60}} \\
                    & \texttt{\textbf{vertex69}} \\
                    & \texttt{\textbf{vertex70}} \\
  \texttt{vertex74} & \texttt{vertex74} \\
  \texttt{vertex75} & \texttt{vertex75} \\
  \texttt{vertex80} & \texttt{vertex80} \\
  \end{tabular}
\end{table}

\subsection{Best standalone CH}

%         NB class BestStandaloneCH

We shall now try to find the best standalone CH, that is the CH that finds on average the best solutions when run without any IHs. We need to set a time limit for \heu{Glpk} to make it an instance of \heu{Truncated Branch \& Bound}, and we shall use 1 second. This is the smallest time limit possible for GLPK and it is still a reasonably short time, fair to other CHs.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Quad Core \\
  Input data        & all official test data sets \\
  Iterations        & 50 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

We will use all the official data sets, set the pool size to 10 where applicable, $\alpha$ and $\beta$ to 1. This experiment will consist of 50 iterations * 11 data sets * 6 CHs = 3300 experimental configurations. See the Listing \ref{listing-experiment-best-standalone-ch} for details. This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.

\begin{algorithm}
\caption{Best Standalone CH set generation}
\label{listing-experiment-best-standalone-ch}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$file \in $ official test data}
	\FOR{$i = 1 \to 50$}
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \emptyset\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Fuzzy}, IH = \emptyset\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Incremental}, IH = \emptyset\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Removal}, IH = \emptyset\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

For data sets \heu{XMA-c}, \heu{XMA-p}, \heu{MSH} and \heu{NTH} every CH found the optimum every time. Graphs representing the results for remaining data sets can be found in Figure \ref{image-experiments-best-standalone-ch}.

\begin{figure}
  \caption{Best Standalone CH}
  \label{image-experiments-best-standalone-ch}
  \centering
  	\subfigure[\dataset{OVA1}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA1}}
  	\subfigure[\dataset{OVA2}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA2}}
  	\subfigure[\dataset{OVA3}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA3}}
  	\subfigure[\dataset{XMD}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-XMD}}
   	\subfigure[\dataset{100-100}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-100}}
  	\subfigure[\dataset{100-200}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-200}}
  	\subfigure[\dataset{100-1000}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-1000}}
\end{figure}

We can see that \heu{Glpk} wins in every single case. We will start from there and try to build upon this result.

\subsection{Best IH for Glpk}

%         NB class BestIHForGlpk

The next logical step will be to try to add one IH after the best CH we have found, \heu{Glpk}. We will investigate all IHs except for \heu{RandomRemove} and \heu{RemoveWorst}, which cannot help us at this time.

We should note that the combination best CH - best IH found this way does not necessarily need to be the best one, because we find it using a hungry approach.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
  Iterations        & 50 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

This experimental set will contain 6 data sets * 50 iterations * 4 IHs = 1200 experimental configurations. Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an overkill for easier data sets.

\begin{algorithm}
\caption{Best IH for \heu{Glpk} set generation}
\label{listing-experiment-best-ih-for-glpk}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
	\FOR{$i = 1 \to 50$}
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Crossover}(ratio = 0.1, limit = 1)\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Hungry}\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Local Branching}(ratio = 0.1, limit = 1)\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

The results are listed in Table \ref{table-experiments-best-ih-for-glpk}. We shall denote \textit{improvement} the absolute increase in quality after running \heu{Glpk} and after running the IH. The table now lists for each data set and each IH the average improvement as well as the standard deviation of the improvement. Bold number represents the best IH for that specific data set. \heu{Mutation} proves to be the best IH for 3 out of 6 data sets.

\begin{table}
  \caption{Best IH for \heu{Glpk}}
  \bigskip
  \label{table-experiments-best-ih-for-glpk}
  \centering
  \begin{tabular}{l || l | l || l | l}
    Data set & \heu{Hungry} & \heu{Hungry} & \heu{Crossover} & \heu{Crossover} \\
     & improvement & improvement & improvement & improvement \\
     & avg & stdev & avg & stdev \\
    \hline
    \dataset{80-320} & 0.00017 & 0.00118 & 0.00017 & 0.00118 \\
    \dataset{90-405} & 0.00502 & 0.00618 & 0.00033 & 0.00165 \\
    \dataset{100-500} & 0.00664 & 0.00667 & 0.00016 & 0.00081 \\
    \hline
    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
    \dataset{100-1000} & 0.01630 & 0.01294 & 0.00180 & 0.00506 \\
  \end{tabular}    
	% TODO separate somehow
  \begin{tabular}{l || l | l || l | l}
    Data set & \heu{LB} & \heu{LB} & \heu{Mutation} & \heu{Mutation} \\
     & improvement & improvement & improvement & improvement \\
     & avg & stdev & avg & stdev \\
    \hline
    \dataset{80-320} & \textbf{0.00072} & 0.00223 & 0.00064 & 0.00218 \\
    \dataset{90-405} & 0.00698 & 0.00616 & \textbf{0.00851} & 0.00659 \\
    \dataset{100-500} & 0.00796 & 0.00797 & \textbf{0.00964} & 0.00804 \\
    \hline
    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
    \dataset{100-1000} & 0.01710 & 0.01188 & \textbf{0.02337} & 0.01558 \\
  \end{tabular}
\end{table}

\subsubsection{\heu{Random} as CH}

%         NB class CHForMutation

TODO with \heu{Mutation} from the previous step, how about using Random as CH? Do we get an improvement?

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
  Iterations        & 50 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

\begin{algorithm}
\caption{\heu{Random} as CH set generation}
\label{listing-experiment-ch-for-mutation}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
	\FOR{$i = 1 \to 50$}
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{figure}
  \caption{CH for \heu{Mutation}}
  \label{image-experiment-ch-for-mutation}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/ch-for-mutation}
\end{figure}

TODO odd are random, even are Glpk

TODO interpretation - Glpk as CH takes always one second, with simpler inputs almost always finds the optimum (thus the capping), with more complex is Random not that bad in the end

\subsection{Various $\alpha$, $\beta$}

TODO remind what these parameters mean, remind that because these are weights, only their ratio matters. We have thus a 1-dimensional problem. However, for simplicity's sake we will use 25 combinations of alpha and beta and normalize them only during evaluation.

TODO we do not expect any changes in performance and will limit the inquiry to different ID sets produced under different settings

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & all official test data sets \\
  Iterations        & 1 \\
  Pool size         & 1 \\
  $\alpha$, $\beta$ & $\{0.1, 0.25, 0.5, 0.75, 1\} \times \{0.1, 0.25, 0.5, 0.75, 1\}$ \\ % TODO this is not the best way to write it
  \hline
\end{tabular}
\bigskip
\end{center}

\begin{algorithm}
\caption{Various values of $\alpha$ and $\beta$ set generation}
\label{listing-experiment-various-betas}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$\alpha \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
  \FOR{$\beta \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
    \FOR{$file \in $ official test data}
    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1, alpha = \alpha, beta = \beta), IH = \emptyset\}$
    \ENDFOR
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{100-1000}, \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.

TODO remaining datasets showed various dependencies on alpha and beta, because this question is not central to our interest we shall mention only 2 examples

TODO in 100-200 we have a clear boundary, in OVA1 (representative of OVAs in general) we have unclear boundary

\subsubsection{Results for \dataset{100-200}}

2 different sets (too long to list here), very nice boundary when we normalize alpha and beta

\subsubsection{Results for \dataset{OVA1}}

TODO found 2 ID sets:

\begin{table}
  \caption{Different ID sets found for \dataset{OVA1}}
  \bigskip
  \label{table-experiments-various-betas-ova1}
  \centering
  \begin{tabular}{c || c}
    ID set 1: element@attribute & ID set 2: element@attribute \\
    \hline
    \texttt{aff@fa} & \texttt{aff@fa} \\
    \texttt{com@ty} & \texttt{com@ty} \\
    \texttt{cre@da} & \texttt{cre@da} \\
    \texttt{cri@te} & \texttt{cri@te} \\
    \texttt{cve@st} & \texttt{cve@st} \\
    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} \\ % TODO again, my bold typewriter does not work
    \texttt{fil@co} & \texttt{fil@co} \\
    \texttt{mod@da} & \texttt{mod@da} \\
    \texttt{ova@xs} & \texttt{ova@xs} \\
    \texttt{pat@op} & \texttt{pat@op} \\
    \texttt{sof@op} & \texttt{sof@op} \\
    \texttt{sta@da} & \texttt{sta@da} \\
    \texttt{sub@or} & \texttt{sub@or} \\
    \texttt{sbt@te} & \texttt{sbt@te} \\
  \end{tabular}
\end{table}

\begin{table}
  \caption{Effect of $\alpha$, $\beta$ on ID set found for \dataset{OVA1}}
  \bigskip
  \label{table-experiments-various-betas-ova1-effect}
  \centering
  \begin{tabular}{c | c  c  c  c  c}
    $\alpha$ \textbackslash $\beta$ & 0.1 & 0.25 & 0.5 & 0.75 & 1 \\
    \hline
    0.1  & 1 & 2 & 2 & 1 & 1 \\
    0.25 & 2 & 2 & 2 & 1 & 2 \\
    0.5  & 2 & 1 & 1 & 1 & 2 \\
    0.75 & 1 & 2 & 1 & 2 & 1 \\
    1    & 1 & 2 & 1 & 1 & 2 \\
  \end{tabular}
\end{table}

TODO linear - we have more than 1 combination that has the same ratio (explain what this ratio is)

\begin{table}
  \caption{Effect of $\alpha$\ ratio on ID set found for \dataset{OVA1}}
  \bigskip
  \label{table-experiments-various-betas-ova1-ratio-effect}
  \centering
  \begin{tabular}{c | c || c |  c}
    $\alpha$ ratio & ID set & $\alpha$ ratio & ID set \\
    \hline
    0,091	& 1 & 0,500	& 2 \\
    0,118	& 1 & 0,500	& 2 \\
    0,167	& 2 & 0,571	& 1 \\
    0,200	& 2 & 0,600	& 1 \\
    0,250	& 1 & 0,667	& 1 \\
    0,286	& 2 & 0,667	& 1 \\
    0,333	& 2 & 0,714	& 2 \\
    0,333	& 2 & 0,750	& 2 \\
    0,400	& 1 & 0,800	& 2 \\
    0,429	& 1 & 0,833	& 2 \\
    0,500	& 1 & 0,882	& 1 \\
    0,500	& 2 & 0,909	& 1 \\
    0,500	& 1 &       &   \\
  \end{tabular}
\end{table}

TODO no clear separation, same ratio producing different sets? Randomness in GLPK input.

\subsubsection{Interpretation}

TODO we just note that there are differences, but will not investigate any further

TODO artificial data always has only one attribute

\subsection{Ignoring text data}

%         NB class IgnoreTextData

TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & \dataset{XMA-p} \\
  Iterations        & 50 \\
  Pool size         & 1 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

\begin{algorithm}
\caption{Ignoring text data set generation}
\label{listing-experiment-ignore-text-data}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\FOR{$i \in 1 \to 50 $}
  \STATE $ES \gets ES \cup \{\dataset{XMA-p}, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
\ENDFOR
\STATE \textbf{set} ``ignore text data"
\FOR{$i \in 1 \to 50 $}
  \STATE $ES \gets ES \cup \{\dataset{XMA-p}, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{figure}
  \caption{Ignoring Text Data}
  \label{image-experiment-ignore-text-data}
  \centering
    \includegraphics[width=\textwidth]{images/experiments/ignore-text-data}
\end{figure}

TODO interpretation - grammar extraction is very similar if not a bit worse, model creation is about 50\% faster.

\subsection{Chaining the IHs}

%         NB class ChainedIHsX

TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.

TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit. First we set all ratios to 0.1, as soon as we find the best scenario, we will tweak it.

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & all official test data sets \\
  Iterations        & 20 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

\begin{algorithm}
\caption{Chaining IHs set generation}
\label{listing-experiment-chaining-ihs}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$
\STATE $\heu{MUT} \gets \heu{Mutation}(ratio = 0.1, limit = 1)$
\STATE $\heu{CX} \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
\STATE $\heu{LB} \gets \heu{LocalBranching}(ratio = 0.1, limit = 1)$
\STATE $\heu{RR} \gets \heu{RandomRemove}(ratio = 0.1)$
\STATE $\heu{H} \gets \heu{Hungry}$
\STATE $\heu{RW} \gets \heu{RemoveWorst}$
\STATE $IHs \gets \emptyset$

\STATE $IHs \gets IHs \cup (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})$

\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RW}, \heu{MUT})$

\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RR}, \heu{MUT}, \heu{RW}, \heu{LB}, \heu{RW}, \heu{H})$

\FOR{$ih \in IHs$}
  \FOR{$file \in $ official test data}
    \FOR{$i = 1 \to 20$}
      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih, limit = 10\}$
    \ENDFOR
  \ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

TODO result graphs - just a \textbf{commented} sample

\dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p} boring - finds the optimum in the first step

TODO choose the best - see how many times in the 20 runs did they find optimum

\begin{table}
  \caption{Performance of various IH chains}
  \bigskip
  \label{table-experiments-chained-ihs-tweaking}
  \centering
  \begin{tabular}{l || c | c | c}
    Dataset & Strategy 1 & Strategy 2 & Strategy 3 \\
    \hline
    \dataset{100-100}  & \textbf{20} & 19 & \textbf{20} \\
    \dataset{100-200}  & \textbf{19} & 18 & 17 \\
    \dataset{100-1000} & 4  & 1  & \textbf{5}  \\
    \dataset{OVA1}     & \textbf{20} & \textbf{20} & \textbf{20} \\
    \dataset{OVA2}     & \textbf{19} & 13 & 18 \\
    \dataset{OVA3}     & 17 & 18 & \textbf{20} \\
    \end{tabular}
  \bigskip
  Each cell represents the number of times (out of 20) the strategy found the optimum - more is better.
\end{table}

Strategies 1 and 3 are very similar, we shall choose \textbf{1} for its simplicity to tune the parameters.

\subsubsection{Improving Strategy 1}

%         NB class ChainedIHs1Tweak

TODO remind what Strategy 1 is, list all ratios we will be using

TODO we will be gathering for each run the following: what were the parameters, how long did it take and whether it found optimum

\begin{center}
\bigskip
\begin{tabular}{| l | l |}
  \hline
  \hline
  Machine           & Dual Core \\
  Input data        & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3} \\
  Iterations        & 50 \\
  Pool size         & 10 \\
  $\alpha$, $\beta$ & $1$, $1$ \\
  \hline
\end{tabular}
\bigskip
\end{center}

\begin{algorithm}
\caption{Chained IHs - Tuning set generation}
\label{listing-experiment-chained-ihs-tuning}
\begin{algorithmic}
\ENSURE experimental set $ES$
\STATE $ES \gets \emptyset$

\STATE $RW \gets \heu{RemoveWorst}$

\FOR{$rrRatio \in \{0, 0.05, 0.1, 0.2, 0.5\}$}
\FOR{$mutRatio \in \{0.05, 0.1, 0.2\}$}
\FOR{$cxRatio \in \{0.05, 0.1, 0.2\}$}
  \STATE $RR \gets \heu{RandomRemoval}(ratio = rrRatio)$
  \STATE $MUT \gets \heu{Mutation}(ratio = mutRatio, limit = 1)$
  \STATE $CX \gets \heu{Crossover}(ratio = cxRatio, limit = 1)$
  \FOR{$file \in \{\dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3}\}$}
    \FOR{$i = 1 \to 50$}
      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})\}$
    \ENDFOR
  \ENDFOR
\ENDFOR
\ENDFOR
\ENDFOR
\RETURN $ES$
\end{algorithmic}
\end{algorithm}

\begin{table}
  \caption{Performance of Strategy 1 depending on parameters - excerpt}
  \bigskip
  \label{table-experiments-chained-ihs-tweaking}
  \centering
  \begin{tabular}{c | c | c || c | c | c | c | c | c}
    \heu{RR} & \heu{MUT} & \heu{CX} & \dataset{100-100} & \dataset{100-1000} & \dataset{100-200} & \dataset{OVA1} & \dataset{OVA2} & \dataset{OVA3} \\
    \hline
    0.2	& 0.2	& 0.2	   & 2749.58	& \textbf{8633.76}	  & 3029.82	& 110.82	& 98.68	& 2413.86 \\
    0.5	& 0.05	& 0.05 & 1031.24	& 10324.46	& 1371.08	& \textbf{71.88}	  & 63.86	& 1282.94 \\
    0.5	& 0.05	& 0.2	 & \textbf{919.74}	  & 10978.58	& 1322.02	& 74.46	  & 62.42	& \textbf{1217.00} \\
    0.5	& 0.1	& 0.2	   & 1002.20	& 11030.64	& \textbf{1288.06}	& 85.22	  & 80.96	& 1480.90 \\
    0.5	& 0.2	& 0.1	   & 1674.78	& 9779.82	  & 2007.42	& 108.14	& \textbf{58.56}	& 1827.46 \\
    \end{tabular}
  \bigskip
  First 3 columns are ratios for specified heuristics. Remaining columns are average run times in ms. Bold are the best times in that column.
\end{table}

TODO interpretation - we calculated average runtimes for each combination of ratios and each file. Then we picked the ratio combination that yielded the lowest average runtimes.

RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)

TODO compare this to pure GLPK run - for example for \dataset{100-1000}, we got from TODO (90+ on average) seconds to on average TODO seconds (when the 10 second limit is lifted).

TODO this will need 2 boxplots again...

\section{The "Best" Algorithm}

TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".

First of all: if we have the time, it is best to just let the GLPK run.
We will find the optimum in this way.
And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.

Second: if we don't have enough time, we should do Strategy 1. In under 10 seconds we will have very good results (often the optimum).

% TODO if possible, write down in lessons learned:
% - always allow for interruptions
% - in experiments, iterations should be the outermost loop
% - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)