### 2011-09-25 doc/theses/vitasek/chapter-2-research.tex

\chapter{Research}

According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set (Section \ref{section-definitions-id-set}) with the maximal weight (Section \ref{section-definitions-weight}) can be transformed into the maximum weighted independent set (Section \ref{section-definitions-is}) and is thus NP-hard.

TODO describe the transformation ID -> IS (other way around is in fidax)

The article suggests a heuristic approach described in Section \ref{section-mip-fidax}, which was incorporated into the framework proposed by this work.

To the best of our knowledge, there are no other articles dealing with this problem.\\

TODO talk about Fajt \cite{fajt} - but that's different, that's keys. Sometimes, it is translatable, but not always.\\

TODO our problem now is to solve max IS.

TODO other approaches to max IS - approximation methods?