### 2011-10-09 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress

\chapter{MIP Approach}
\label{chapter-mip}

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by applying heuristical approaches to improve the performance of the process.

\section{ID Set to IS Formulation}

Given $C = \{m_1, \ldots, m_n\}$ a set of all AMs in a document, we construct a graph $G = (V,E)$ as follows. For each AM $m_i \in C$ we create a vertex $v_{name(m_i)}$. Two vertices $v_{name(m_i)}$ and $v_{name(m_j)}$ shall be connected by an edge iff they cannot share the same ID set, either because they have the same type ($\tau(m_i) = \tau(m_j)$), or their images intersect ($\iota(m_i) \cap \iota(m_j) \neq \emptyset$). Weight of a vertex $v_{name(m_i)}$ is the weight of the attribute mapping: $w(v_{name(m_i)}) = weight(m_i)$.

Now finding the maximum weighted IS in $G$ finds the maximum (optimal) ID set in the original document.

\section{IS to MIP Formulation}

Given a graph $G = (V,E)$ with a weight function $w: V \rightarrow \mathbb{R}$, we introduce a binary variable $x_i$ for each vertex $v_i \in V$ and an inequality constraint $x_i + x_j \leq 1$ for each edge $e = (v_i, v_j) \in E$. Furthermore we introduce an objective function in form $\sum_{x_i} x_i w(v_i)$.

It is obvious that the objective function and all the constraints consitute a MIP instance, and that solving it finds the maximum weigthed IS in $G$.

\section{Finding ID Sets With GLPK}

By chaining these two translations we can create a MIP formulation for a given set of AMs from a document. Solving this MIP instance will give us the optimal ID set for this document.

GLPK \cite{glpk} is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} (see \cite{land60a} for details) for MIP. An advantage of using \textit{Branch \& Bound} is that while traversing the \textit{search tree} it finds intermittent, sub-optimal solutions. It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\

We will now demonstrate the full process of finding the optimal ID set of an example XML file using GLPK.

\subsection{Example}

Consider again our XML file fragment.

\begin{verbatim}
<x>
  <y a="1" b="2"/>
  <y a="3" c="4"/>
  <y/>
  <z a="1"/>
</x>
\end{verbatim}

Recall that attribute mappings in this example are $C = \{M_{y}^{a}, M_{y}^{b}, M_{y}^{c}, M_{z}^{a}\}$. Corresponding vertices in the IS formulation will be $V = \{v_{y-a}, v_{y-b}, v_{y-c}, v_{z-a}\}$. Edges in the IS formulation will be the following.

\begin{eqnarray*}
(v_{y-a},v_{y-b}) \\
(v_{y-a},v_{y-c}) \\
(v_{y-b},v_{y-c}) \\
(v_{y-a},v_{z-a}) \\
\end{eqnarray*}

First three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.

\begin{figure}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
  \centering
	\includegraphics[width=.25\textwidth]{images/is-representation}
\end{figure}

The next step is the MIP formulation. We do not need to translate from the IS formulation, as the translation from ID set formulation is straightforward, too. For each AM $m$ there will be one binary variable $x_{name(m)}$. Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings. For each pair of AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$. $\mathbf{b}$ will be a vector of ones with corresponding length.

\[
\mathbf{x} =
\begin{bmatrix}
x_{y-a} \\
x_{y-b} \\
x_{y-c} \\
x_{z-a} \\
\end{bmatrix},
\mathbf{c} = 
\begin{bmatrix}
weight(M_{y}^{a}) \\
weight(M_{y}^{b}) \\
weight(M_{y}^{c}) \\
weight(M_{z}^{a}) \\
\end{bmatrix} =
\begin{bmatrix}
0.2 \\
0.2 \\
0.6 \\
0.4 \\
\end{bmatrix},
\mathbf{b} =
\begin{bmatrix}
1 \\
1 \\
1 \\
1 \\
\end{bmatrix},
A =
\begin{pmatrix}
0 & 1 & 1 & 0 \\
1 & 0 & 1 & 0 \\
1 & 1 & 0 & 0 \\
1 & 0 & 0 & 1 \\
\end{pmatrix}
\]

The problem now is, recall, to solve the following.

\begin{eqnarray*}
\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
\end{eqnarray*}

In GLPK \textit{MathProg} language, this translates to the following formulation. % TODO cite mathprog

\begin{scriptsize}
\begin{verbatim}
set AMs;
param Weight {i in AMs};
var x {i in AMs} binary;
maximize z: sum {i in AMs} x[i] * Weight[i];
s.t. c1: x['y-a'] + x['y-b'] <= 1;
s.t. c2: x['y-a'] + x['y-c'] <= 1;
s.t. c3: x['y-b'] + x['y-c'] <= 1;
s.t. c4: x['y-a'] + x['z-a'] <= 1;
data;
set AMs := y-a y-b y-c z-a;
param Weight :=
y-a 0.6
y-b 0.2
y-c 0.2
z-a 0.4;
end;
\end{verbatim}
\end{scriptsize}

We can use this as an input for the GLPK solver, and we get the solution.

\begin{scriptsize}
\begin{verbatim}
...
Problem:    glpk_input
Rows:       5
Columns:    4 (4 integer, 4 binary)
Non-zeros:  12
Status:     INTEGER OPTIMAL
Objective:  z = 0.6 (MAXimum)

...

   No. Column name       Activity     Lower bound   Upper bound
------ ------------    ------------- ------------- -------------
     1 x[y-a]       *              1             0             1 
     2 x[y-b]       *              0             0             1 
     3 x[y-c]       *              0             0             1 
     4 x[z-a]       *              0             0             1

...
\end{verbatim}
\end{scriptsize}

This output tells us that the solution is $x_{y-a} = 1$, $x_{y-b} = 0$, $x_{y-c} = 0$ and $x_{z-a} = 0$. This means that the optimal ID set with maximum weight contains only the $M_{y}^{a}$ attribute mapping.\\

It is obvious that this approach works and for any possible input we can let GLPK find the optimal solution. However, sometimes it takes too long to find the optimum (see e.g. Section \ref{section-glpk-comparison}), we should try to improve this process.

\section{Heuristics}

TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)

TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)

TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight

TODO we will be using a pool - what is a pool

TODO define incumbent solution

\begin{figure}
  \caption{Metaheuristic schema}
  \label{image-metaheuristic}
  \centering
    \includegraphics[width=\textwidth]{images/metaheuristic}
\end{figure}

We will now introduce several heurisitics we have implemented to aid us in search for optimal ID set.

\subsection{Constructions Heuristics}
% TODO can I format things in nomenclature?
\nomenclature{CH}{Construction Heuristic} % TODO decide on how to format CH/IH

When we start to solve a problem using a metaheuristic approach, at first we have no solutions at all. The purpose of a construction heuristic (CH) is then to provide us with at least some solution. This may or may not be already the optimum, in the latter case it will be improved on later using improvement heuristics (IH). Some IHs can profit from a pool of several sub-optimal solutions, and some CHs can produce this pool from them.

\subsubsection{\heu{FIDAX}}
\label{section-mip-fidax}

The first construction heuristic is the algorithm described in \cite{fidax} (we shall call it \heu{FIDAX} from now on). It can trivially be used to give us one feasible solution - it is, after all, a deterministic heuristic algorithm.

The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{listing-ch-fidax}.

\begin{algorithm}
\caption{\heu{FIDAX} CH}
\label{listing-ch-fidax}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets C$ sorted by decreasing size
\STATE Compute the weight $w(m)$ of each $m$ in $C$
\FORALL{$t$ in $\Sigma^E$}
  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
  \STATE Remove from $C'$ all mappings of type $t$ except $m$
\ENDFOR
\FORALL{$m$ in $C'$}
  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
  \IF{$w(m) > \sum_{p \in S} w(p)$}
    \STATE remove all $p \in S$ from $C'$
  \ELSE
    \STATE remove $m$ from $C'$
  \ENDIF
\ENDFOR
\RETURN $C'$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Random}}
\label{heu-ch-random}

One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution. This is obviously a hungry heuristic.

The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free. As we will see later in the experiments (Section \ref{section-experiments-random-fuzzy-fidax}), it performs surprisingly well.

See the Listing \ref{listing-ch-random} for its pseudocode.

\begin{algorithm}
\caption{\heu{Random} CH}
\label{listing-ch-random}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$}
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \WHILE{$s$ is a feasible ID set}
    \STATE $a \gets $ pick at random from $C \backslash S$
    \STATE $s \gets s \cup a$
  \ENDWHILE
  \STATE $r \gets r \cup s$
\ENDFOR
\RETURN $r$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Fuzzy}}
\label{heu-ch-fuzzy}

\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is selected based on \textit{weighted} instead of uniform random. The weight used here is the usual weight of an AM as defined in Section \ref{section-definitions-weight}. Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.

This is again a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.

% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?

\begin{algorithm}
\caption{\heu{Fuzzy} CH}
\label{listing-ch-fuzzy}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$}
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \STATE $C' \gets C$

  \WHILE{$C' $ not empty}
    \STATE $a \gets $ pick at weighted random from $C'$
    \IF{$s \cup a$ is a feasible ID set}
      \STATE $s \gets s \cup a$
      \STATE $C' \gets C' \backslash a$
    \ENDIF
    \FORALL{$c \in C'$}
      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
        \STATE $C' \gets C' \backslash c$
      \ENDIF
    \ENDFOR
  \ENDWHILE

  \STATE $r \gets r + s$
\ENDFOR
\RETURN $r$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Incremental}}

This trivial heuristics sorts all candidate AMs by their decreasing weights (Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.

See Listing \ref{listing-ch-incremental} for its pseudocode.

\begin{algorithm}
\caption{\heu{Incremental} CH}
\label{listing-ch-incremental}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by decreasing weight
\STATE $s \gets $ empty solution
\FORALL{$c \in C'$}
  \IF{$s \cup c$ is a feasible ID set}
    \STATE $s \gets s + c$
  \ENDIF
\ENDFOR
\RETURN $s$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Removal}}

This is basically a reversal of the idea from the \heu{Incremental} heuristic - start with a solution containing all the candidate AMs. This probably does not satisfy the ID set condition. Therefore, sort them by increasing size and start removing them from the solution, until it satisfies the ID set condition. Again, this is a hungry heuristic returning only one solution.

See Listing \ref{listing-ch-removal} for its pseudocode.

\begin{algorithm}
\caption{\heu{Removal} CH}
\label{listing-ch-removal}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by increasing weight
\STATE $s \gets C'$
\FORALL{$c \in s$}
  \IF{$s$ is a feasible ID set}
    \RETURN $s$
  \ENDIF
  \STATE $s \gets s \backslash c$
\ENDFOR
\end{algorithmic}
\end{algorithm}

\subsubsection{Truncated Branch \& Bound - \heu{Glpk}}

This construction heuristic will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK. Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found. Limiting the run time gives us the best solution found so far, which means this is a construction heuristic.

To be able to create a pool of solutions, the GLPK input in MathProg langauge is always randomly shuffled by changing the order in which variables and constraints appear. This is causes the solver to explore the search tree in various orders, producing different solution in each of the time-constrained runs.

\subsection{Improvement Heuristics}
\label{section-mip-ihs}

\nomenclature{IH}{Improvement Heuristic}

Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in the end. We will need two notions to describe their behavior.

\paragraph{Intensification}

is the attempt to move the solution towards the nearby local optimum in the solution space.

\paragraph{Diversification}

is the attempt to move the solution away (escape) from from the local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\

A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum.

\subsubsection{\heu{Identity}}

This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.

\begin{algorithm}
\caption{\heu{Identity} IH}
\label{listing-ih-identity}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE the same pool of feasible solutions
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Remove Worst}}

This trivial IH tries to improve the solution pool by removing the worst solution (i.e. the one with the lowest quality). This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.

See Listing \ref{listing-ih-removeworst} for details.

\begin{algorithm}
\caption{\heu{Remove Worst} IH}
\label{listing-ih-removeworst}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE pool of feasible solutions
\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
\RETURN $FP \backslash s_{min}$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Random Remove}}

This is again a rather trivial diversification improvement heuristic. By removing a random subset of specified size from each solution in the pool, it provides the variability needed to escape from local optima. % TODO make sure we have talked about local optima before

The number of AMs to remove from each solution is specified as ratio from $(0, 1)$ of the solution size (number of AMs in solution). For example, \heu{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).

This heuristic returns a pool of solutions of the same size as it got on its input.

See Listing \ref{listing-ih-randomremove} for pseudocode.

\begin{algorithm}
\caption{\heu{Random Remove} IH}
\label{listing-ih-randomremove}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE $K \gets k * |s|$
  \STATE remove $K$ random AMs from $s$
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Hungry}}

This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.

\heu{Hungry} tries to improve each solution in the feasible pool in the following way. It orders all candidate AMs not present in the solution by decreasing weight. Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition. The resulting solution (whether any AMs were added or not) is then returned to the pool. This is then intensification, and Listing \ref{listing-ih-hungry} captures the process.

\begin{algorithm}
\caption{\heu{Hungry} IH}
\label{listing-ih-hungry}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE \COMMENT {improve a single solution}
  \STATE $C' \gets C \backslash s$
  \STATE $C' \gets C'$ sorted by decreasing weight
  \FORALL{$c \in C'$}
    \IF{$s \cup c$ is a feasible ID set}
      \STATE $s \gets s \cup c$
    \ENDIF
  \ENDFOR
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Mutation}}

\heu{Mutation} is based on the following idea. We assume that an incumbent solution may already contain some AMs belonging to the optimal solution. We will take a random guess and fix some of these AMs: add new constraints to the MIP formulation fixing values of the respective variables to 1.

This new formulation contains less free variables and should be easier to solve, probably even to optimum. We run GLPK again using this constrained formulation, enforcing again a time limit. Solution found this way is a feasible solution of the original problem, however the optimum is not necessarily the same as in unconstrained formulation. It is an intensification approach: we limit the search to the neighborhood of an already found solution.\\

\heu{Mutation} changes the MIP formulation in following way. For every AM $AM_F$ fixed to appear in the solution a following constraint is added to GLPK input:
\[s.t. f_{index}: x['name(AM_F)'] = 1;\]
$index$ is a unique integer to number all the constraints.

Additionaly, every other mapping $AM_i$ colliding with $AM_F$ ($\iff \iota(AM_F) \cap \iota(AM_i) \neq \emptyset$) will cause the following constraint to be added:
\[s.t. f_{index}: x['name(AM_i)'] = 0;\]
And the original constraint in form:
\[s.t. c_{index}: x['name(AM_F)'] + x['name(AM_i)'] <= 1;\]
will not be included.\\

Listing \ref{listing-ih-mutation} captures the process of randomly selecting a specified fraction of AMs of the incumbent solution to fix, then running GLPK again. \heu{Mutation} requires pool of at least one solution as input, and adds the improved solution to the result pool.

\begin{algorithm}
\caption{\heu{Mutation} IH}
\label{listing-ih-mutation}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ fraction of AMs to fix
\ENSURE pool of feasible solutions
\STATE $incumbent \gets $ incumbent solution in $FP$
\STATE $K \gets k * |incumbent|$
\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Crossover}}

TODO explain how this works and link wise books

TODO explain how this translates to GLPK input - it's again simple fixing to 1, but we get the list of AMs in a different manner.

See listing \ref{listing-ih-crossover}.

\begin{algorithm}
\caption{\heu{Crossover} IH}
\label{listing-ih-crossover}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of solutions to look for commonalities in
\ENSURE pool of feasible solutions
\STATE $K \gets k * |FP|$
\STATE $FP' \gets K$ random solutions $\in FP$
\STATE $am \gets$ AMs found in all solutions $\in FP'$
\STATE fix $am$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

\subsubsection{\heu{Local Branching}}

TODO explain how this works and link wise books

TODO explain how this translates to GLPK input

A new constrain describing the maximal allowed distance from the incumbent solution is added to GLPK input.

\[s.t. LB: sum\{i\ in\ INCUMBENT\} (1 - x[i]) + sum\{i\ in\ REMAINING\} x[i] \leq k;\]

Where $INCUMBENT$ is a set of names of AMs in the incumbent solution, $REMAINING$ is a set of all the AMs not included in the incumbent solution and $k$ is the requested maximal distance.

See Listing \ref{listing-ih-localbranching} for pseudocode.\\

\begin{algorithm}
\caption{\heu{Local Branching} IH}
\label{listing-ih-localbranching}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
\ENSURE pool of feasible solutions
\STATE $K \gets k * |$total AM count$|$
\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
\STATE add max Hamming distance requirement to GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

TODO mention that combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} is a sort of genetic algorithm

\section{IDREF}

Once an ID set is found, regardless of how exactly, it is easy to find the IDREF set, i.e. the attribute mappings that can be declared as IDREF. % TODO link FIDAX, they did it first.

First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set. This is because the specification % TODO link to the specific anchor in the page
does not allow an attribute to be ID and IDREF (IDREFS) at the same time. Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}). % TODO link candidate AMs

Second, find the image of the ID set as the union of images of all the AMs in this ID set.

\[\iota(ID) = \bigcup_{m \in ID} \iota(m)\]

Now the IDREF set contains all the AMs whose images are a subset of the ID set image.

\[\iota(c) \subset \iota(ID) \Rightarrow c \in IDREF\]

This can be easily determined in a loop over the list of candidates. The process is captured in Listing \ref{listing-idref}.

\begin{algorithm}
\caption{IDREF Search}
\label{listing-idref}
\begin{algorithmic}
\REQUIRE $AMs$ list of all AMs
\REQUIRE $ID$ ID set as a list of AMs
\ENSURE $IDREF$ set as a list of AMs
\STATE $IDREF \gets \emptyset$
\STATE $candidates \gets AMs \backslash ID$
\STATE $\iota(ID) \gets \bigcup_{m \in ID} \iota(m)$
\FORALL{$c \in candidates$}
  \IF{$\iota(c) \subset \iota(ID)$}
    \STATE $IDREF \gets IDREF \cup c$
  \ENDIF
\ENDFOR
\RETURN $IDREF$
\end{algorithmic}
\end{algorithm}