### 2011-10-27 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

\documentclass[a4paper,12pt,oneside]{report}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
\usepackage{xstring}
\usepackage{wrapfig}
\usepackage{appendix}
\usepackage{nomencl}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multicol}
\usepackage[pdfborder=0 0 0]{hyperref}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{complexity}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\heu}[1]{\texttt{\textit{#1}}}
\newcommand{\dataset}[1]{\texttt{\textit{#1}}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
\linespread{1.5}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\renewcommand{\algorithmicforall}{\textbf{for each}}
\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}

\renewcommand{\nomname}{List of Abbreviations}
\makenomenclature

\theoremstyle{definition}
\newtheorem{define}{Definition}[chapter]

% TODO uncomment this to get much shorter ToC
% \setcounter{tocdepth}{1}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{4cm}
\setlength{\evensidemargin}{4cm}
\setlength{\topmargin}{25mm}
\setlength{\footskip}{1cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textheight}{23.7cm}
\setlength{\textwidth}{14.5cm}
\let\openright=\clearpage

\def\mfauthor{Matej Vitásek}
\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá Ph.D.}
\def\mfplacedate{Praha, 2011}

% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
\makeatletter
\def\@makechapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries \thechapter. #1
   \par\nobreak
   \vskip 20\p@
}}
\def\@makeschapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries #1
   \par\nobreak
   \vskip 20\p@
}}
\makeatother

\def\chapwithtoc#1{
\chapter*{#1}
\addcontentsline{toc}{chapter}{#1}
}

\begin{document}

\lefthyphenmin=2
\righthyphenmin=2

% -----------------------------------------------------------------------------

%                            FINAL CLEANUP

% - make sure all chapter names are in First Caps
% - make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
% - get rid of all short forms: won't -> will not
% - make sure there are no singletons
% - make sure LaTeX does not complaint much about under/overfullness
% - make sure everything is defined and referenced
% - make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
% - make sure that when referring to heus in algorithms, e.g. when describing how experimental sets are constructed, we correctly refer to their parameters as described in their pseudocodes
% - the official name for things like OVA1 and 100-100 is (test) data set. Make sure I am not calling it files or anything else.
% - make sure labels have consistent names
% - ratio -> FRACTION of XYZ to fix/remove/...
% - what style of quotation marks to use?

%%% Titulní strana práce ======================================================
\pagestyle{empty}
\begin{center}

\large

Charles University in Prague

\medskip

Faculty of Mathematics and Physics

\vfill

{\bf\Large MASTER THESIS}

\vfill

\centerline{\mbox{\includegraphics[width=60mm]{logo}}}

\vfill
\vspace{5mm}

{\LARGE \mfauthor}

\vspace{15mm}

% exactly as assigned
{\LARGE\bfseries Inference of XML Integrity Constraints}

\vfill

% Název katedry nebo ústavu, kde byla práce oficiálně zadána
% (dle Organizační struktury MFF UK)
Department of Software Engineering

\vfill

\begin{tabular}{rl}

Supervisor of the master thesis: & 	\mfadvisor \\
\noalign{\vspace{2mm}}
Study programme: & Informatika \\
\noalign{\vspace{2mm}}
Specialization: & ISS \\
\end{tabular}

\vfill

% Zde doplňte rok
\mfplacedate

\end{center}

\newpage % ============================================================
%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
\openright

\noindent
I would like to thank everyone participating in the jInfer project for creating a~solid foundation for this work, namely Irena Mlýn\-ko\-vá, Mi\-chal Klem\-pa, Má\-rio Mi\-ku\-la, Ro\-bert Sme\-ta\-na and Mi\-chal Švi\-rec. The thanks goes again to Ms. Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others.

\newpage % ============================================================
%%% Strana s čestným prohlášením k diplomové práci
\vglue 0pt plus 1fill
\noindent
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

\medskip\noindent
I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm}

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
signature
\hss}}

\vspace{20mm}

\newpage % ============================================================
%%% Povinná informační strana diplomové práce
\vbox to 0.5\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?

Autor: \mfauthor

Katedra:  Katedra softwarového inženýrství

Vedoucí diplomové práce: \mfadvisor{} % TODO does this have to be here?, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství

Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schému existujících XML dokumentů. Jelikož je odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se na integritní omezení. Několik jich popisujeme, pozornost pak soustředíme na ID/\.IDREF/\.IDREFS atributy z DTD. Na bázi článku od Barbosa a Menelzon stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.

Klíčová slova:  XML, ID atributy, odvozování

\vss}\nobreak\vbox to 0.49\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Title:  Inference of XML Integrity Constraints

Author: \mfauthor

Department: Department of Software Engineering

Supervisor: \mfadvisor{} % TODO does this have to be here?, Department of Software Engineering

Abstract: In this work we expand upon the previous efforts to extract (infer) schema information from existing XML documents. We find the inference of \textit{structure} to be sufficiently researched and focus further on integrity constraints. After briefly introducing some of them we turn our attention to ID/\.IDREF/\.IDREFS attributes in DTD. Building on the research by Barbosa and Menelzon we introduce a heuristic approach to the problem of finding optimal ID set. The approach is evaluated and tuned in a wide range of experiments.

Keywords: XML, ID attributes, inference

\vss}

\newpage

%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
%%% na začátku práce, místo na jejím konci.

\openright
\pagestyle{plain}
\setcounter{page}{1}
\tableofcontents

% =============== TEXT ================
\newpage

\chapwithtoc{Preface}

Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data. However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...). To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer. But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well. One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).

This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.

TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
We will focus on the first kind - ID and IDREF attributes.

TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.

\section{Structure of the thesis}

The thesis will be structured as follows.

First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.

Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.

This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.

We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".

Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.

An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.

\section{Conventions}

As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{get\-Heu\-ris\-tic()}. Names of specific heuristics will be written like this: \heu{Mutation}. Name of test data sets will be written like this: \dataset{OVA1}.\\

TODO perhaps explain the way we will be writing pseudocode?\\

There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.\\

TODO we will be ignoring O() complexities

\include{chapter-1-definitions}

\include{chapter-2-research}

\include{chapter-3-mip}

\include{chapter-4-1-experiments-setup}

\include{chapter-4-2-experiments-results}

\include{chapter-5-future}

\include{chapter-6-conclusion}

%%% Seznam použité literatury
\newpage
\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}

\listofalgorithms
\addcontentsline{toc}{chapter}{List of Algorithms}

%%% Tabulky v diplomové práci, existují-li.
%\chapwithtoc{List of Tables}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.

\printnomenclature[2cm]
\addcontentsline{toc}{chapter}{List of Abbreviations}
\label{chapter-list-abbreviations}

%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
%%% textu práce. Přílohy se číslují.

\appendix

\openright
\addcontentsline{toc}{chapter}{Appendices}

\include{appendix-a-jinfer}

\include{appendix-b-iss}

\include{appendix-c-trace}

% TODO a QED square

\end{document}