### 2011-11-01 doc/theses/mikula/TODO_NAZOV.tex
Master thesis: In progress.

\chapter{Diskusia k problemom}
\todo[inline]{Vhodny nazov pre tuto kapitolu}

1) odvodzovanie struktury z queries

2) pattern searching vs lexical, syntax and semantics analysis