### 2011-11-13 doc/theses/vitasek/diplomka.tex

\documentclass[a4paper,12pt,twoside]{report}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
\usepackage{xstring}
\usepackage{wrapfig}
\usepackage[title,titletoc]{appendix}
%\usepackage{textgreek}
\usepackage{nomencl}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multicol}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{complexity}
%\usepackage{refcheck}
\usepackage[pdfborder=0 0 0,pdftex,unicode]{hyperref}
\DeclareFontShape{OT1}{cmtt}{bx}{n}{<5><6><7><8><9><10><10.95><12><14.4><17.28><20.74><24.88>cmttb10}{}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\heu}[1]{\texttt{\textit{#1}}}
\newcommand{\dataset}[1]{\texttt{\textit{#1}}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
\linespread{1.5}
%\onehalfspacing

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\renewcommand{\algorithmicforall}{\textbf{for each}}
\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}

\renewcommand{\nomname}{List of Abbreviations}
\makenomenclature

\theoremstyle{definition}
\newtheorem{define}{Definition}[chapter]

% TODO uncomment this to get much shorter ToC
% \setcounter{tocdepth}{1}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{4cm}
\setlength{\evensidemargin}{4cm}
\setlength{\topmargin}{25mm}
\setlength{\footskip}{1cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textheight}{23.7cm}
\setlength{\textwidth}{14.5cm}
\let\openright=\clearpage

\def\mfauthor{Matej Vitásek}
\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá Ph.D.}
\def\mfplacedate{Praha, 2011}

% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
\makeatletter
\def\@makechapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries \thechapter. #1
   \par\nobreak
   \vskip 20\p@
}}
\def\@makeschapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries #1
   \par\nobreak
   \vskip 20\p@
}}
\makeatother

\def\chapwithtoc#1{
  \chapter*{#1}
  \addcontentsline{toc}{chapter}{#1}
}

\begin{document}

\lefthyphenmin=2
\righthyphenmin=2

% -----------------------------------------------------------------------------

%                            FINAL CLEANUP

% + make sure all chapter names are in First Caps
% + make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
% + get rid of all short forms: won't -> will not
% - make sure there are no singletons
% - make sure LaTeX does not complaint much about under/overfullness
% + make sure everything is defined and referenced
% + make sure there is nothing in the bibliography that is unreferenced
% + make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
% - make sure that when referring to heus in algorithms, e.g. when describing how experimental sets are constructed, we correctly refer to their parameters as described in their pseudocodes
% - the official name for things like OVA1 and 100-100 is (test) data set. Make sure I am not calling it files or anything else.
% - make sure labels have consistent names
% + ratio -> FRACTION of XYZ to fix/remove/...
% + what style of quotation marks to use? (`` and '')

%%% Titulní strana práce ======================================================
\pagestyle{empty}
\begin{center}

\large

Charles University in Prague

\medskip

Faculty of Mathematics and Physics

\vfill

{\bf\Large MASTER THESIS}

\vfill

\centerline{\mbox{\includegraphics[width=60mm]{logo}}}

\vfill
\vspace{5mm}

{\LARGE \mfauthor}

\vspace{15mm}

% exactly as assigned
{\LARGE\bfseries Inference of XML Integrity Constraints}

\vfill

% Název katedry nebo ústavu, kde byla práce oficiálně zadána
% (dle Organizační struktury MFF UK)
Department of Software Engineering

\vfill

\begin{tabular}{rl}

Supervisor of the master thesis: & 	\mfadvisor \\
\noalign{\vspace{2mm}}
Study programme: & Informatika \\
\noalign{\vspace{2mm}}
Specialization: & ISS \\
\end{tabular}

\vfill

% Zde doplňte rok
\mfplacedate

\end{center}

\newpage % ============================================================
%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
\openright

\noindent
I would like to thank everyone participating in the jInfer project for creating a~solid foundation for this work, namely Irena Mlýn\-ko\-vá, Mi\-chal Klem\-pa, Má\-rio Mi\-ku\-la, Ro\-bert Sme\-ta\-na and Mi\-chal Švi\-rec. The thanks goes again to Ms. Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others.

\newpage % ============================================================
%%% Strana s čestným prohlášením k diplomové práci
\vglue 0pt plus 1fill
\noindent
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

\medskip\noindent
I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm}

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
signature
\hss}}

\vspace{20mm}

\newpage % ============================================================
%%% Povinná informační strana diplomové práce
\vbox to 0.5\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?

Autor: \mfauthor

Katedra:  Katedra softwarového inženýrství

Vedoucí diplomové práce: \mfadvisor{} % TODO does this have to be here?, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství

Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schéma existujících XML dokumentů. Jelikož je odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se na integritní omezení. Několik jich popisujeme, pozornost pak soustředíme na ID/\.IDREF/\.IDREFS atributy z DTD. Na bázi článku od Barbosa a Menelzon (2003) stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.

Klíčová slova:  XML, ID atributy, odvozování

\vss}\nobreak\vbox to 0.49\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Title:  Inference of XML Integrity Constraints

Author: \mfauthor

Department: Department of Software Engineering

Supervisor: \mfadvisor{} % TODO does this have to be here?, Department of Software Engineering

Abstract: In this work we~expand upon the previous efforts to~extract (infer) schema information from existing XML documents. We find the~inference of \textit{structure} to~be~sufficiently researched and focus further on~\textit{integrity constraints}. After briefly introducing some of~them we turn our attention to~ID/\.IDREF/\.IDREFS attributes in~DTD. Building on~the~research by~Barbosa and Menelzon (2003) we~introduce a~heuristic approach to~the~problem of~finding optimal ID set. The approach is~evaluated and tuned in~a~wide range of~experiments.

Keywords: XML, ID attributes, inference

\vss}

\newpage

%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
%%% na začátku práce, místo na jejím konci.

\openright
\pagestyle{plain}
\setcounter{page}{1}
\tableofcontents

% =============== TEXT ================
\newpage

  \include{chapter-0-preface}
  \include{chapter-1-definitions}
  \include{chapter-2-research}
  \include{chapter-3-mip}
  \include{chapter-4-1-experiments-setup}
  \include{chapter-4-2-experiments-results}
  \include{chapter-5-future}
  \include{chapter-6-conclusion}

\newpage
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
%\nocite{*}
%\bibliographystyle{alpha}
%\bibliography{literature}
	\include{bibliography}

\newpage
\phantomsection
\addcontentsline{toc}{chapter}{Lists}

\clearpage
\phantomsection
\addcontentsline{toc}{section}{List of Figures}
\listoffigures

\clearpage
\phantomsection
\addcontentsline{toc}{section}{List of Algorithms}
\listofalgorithms

\clearpage
\phantomsection
\addcontentsline{toc}{section}{List of Tables}
\listoftables

\clearpage
\phantomsection
\addcontentsline{toc}{section}{List of Abbreviations}
\label{chapter-list-abbreviations}
\printnomenclature[2cm]

\openright
\clearpage
\phantomsection
\addappheadtotoc
\renewcommand{\appendixname}{Appendix}

\begin{appendices}
  \input{appendix-a-jinfer} % TODO include?
  \include{appendix-b-iss}
  \include{appendix-c-trace}
\end{appendices}

\end{document}