### 2011-11-19 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: merging Lydka round 1.

\chapter{Definitions}
\label{chapter-definitions}

\section{XML Tree}

We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data. Parent nodes are connected to child nodes with the edges. This tree shall be called an \textit{XML tree}. For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.

Without a loss of generality we ignore the actual ordering of nodes in the tree.

\paragraph{Example}

This example introduces an XML file fragment that will be used for demonstration throughout this work. An XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}(a), where each node is annotated with a triple $label(v) : id(v) : value(v)$.

\begin{verbatim}
<x>
  <y a="1" b="2"/>
  <y a="3" c="4"/>
  <y/>
  <z a="1"/>
</x>
\end{verbatim}

\begin{figure}
  \caption{Example XML Tree}
  \label{image-definitions-example-xml-tree} 
  \centering
    \subfigure[XML tree]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree}}
    \subfigure[Attribute Mappings]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree-ams}}
\end{figure}

Furthermore, we denote $\mathcal{I}$ the set of all ids and $\mathcal{V}$ the set of all values in the document. We will need two more definitions from the article.

\begin{define}[Node equality]
	Nodes $v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
\end{define}

\begin{define}[Value equality]
	Nodes $v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
\end{define}

\section{\texttt{ID}, \texttt{IDREF}, \texttt{IDREFS} Attributes}
\label{section-definitions-id-attributes}

According to \cite{Bray:08:EML}, an XML attribute may have the type \texttt{ID}, \texttt{IDREF} or \texttt{IDREFS} (among others). The following constraints are related to these types.

\begin{quote}
\textbf{Validity constraint: \texttt{ID}}

Values of type \texttt{ID} \textsc{must} match the Name production. A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \texttt{ID} values \textsc{must} uniquely identify the elements which bear them.

\textbf{Validity constraint: One \texttt{ID} per Element Type}

An element type \textsc{must not} have more than one \texttt{ID} attribute specified.

\textbf{Validity constraint: \texttt{ID} Attribute Default}

An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED}\footnote{\texttt{\#IMPLIED} means that the attribute has a specified default value.} or \texttt{\#REQUIRED}\footnote{\texttt{\#REQUIRED} means that the attribute cannot be empty.}.

\textbf{Validity constraint: \texttt{IDREF}}

Values of type \texttt{IDREF} \textsc{must} match the Name production, and values of type \texttt{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \texttt{ID} attribute on some element in the XML document; i.e. \texttt{IDREF} values \textsc{must} match the value of some \texttt{ID} attribute.
\end{quote}

\section{Attribute Mappings}
\label{section-definitions-ams}

Now we return to \cite{fidax} to define the notion of an \textit{attribute mapping} (or AM for short). \nomenclature{AM}{Attribute Mapping}
We will use a different definition (without introducing keys from \cite{keX}) that will however give us the same result.

\begin{define}[$\Sigma^E$, $\Sigma^A$, $\Sigma$]
	$\Sigma^E$ is the set of all element labels, $\Sigma^A$ is the set of all attribute labels. $\Sigma = \Sigma^E \cup \Sigma^A$ is their union and effectively the set of all labels in the document.
\end{define}

\begin{define}[Attribute mapping]
	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $\mathcal{I} \times \mathcal{I}$ relation defined by
	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}.\]
\end{define}

Thus the relation $M_{x}^{y}$ contains edges in the XML tree connecting element nodes labeled $x$ and attribute nodes labeled $y$.

We can use projection to retrieve all the unique \textit{ids} of either elements or attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.

\begin{define}[Type of the attribute mapping]
	Attribute mapping $M_{x}^{y}$ is of the \textit{type} $\tau(M_{x}^{y}) = x$.
\end{define}

\paragraph{Example}
The XML tree from Figure \ref{image-definitions-example-xml-tree}(b) has the following non-empty AMs drawn in bold lines: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.

The following example equations hold.
\begin{eqnarray*}
\pi_E(M_{y}^{a}) & = & \{2, 3\} \\
\pi_A(M_{z}^{a}) & = & \{10\} \\
\tau(M_{y}^{c}) & = & y
\end{eqnarray*}

\begin{define}[Image of the attribute mapping]
	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$.
\end{define}

So the image of an AM is a set of all the values of all the attribute nodes contained in the mapping.

\paragraph{Example}
Again referring to the XML tree from Figure \ref{image-definitions-example-xml-tree}, we get the following AM images.
\begin{eqnarray*}
\iota(M_{y}^{a}) & = & \{1, 3\} \\
\iota(M_{y}^{b}) & = & \{2\} \\
\iota(M_{y}^{c}) & = & \{4\} \\
\iota(M_{z}^{a}) & = & \{1\} \\
\end{eqnarray*}

\paragraph{Attribute Mapping Model}
An attribute mapping model is a data structure containing the information about all the AMs in a document, together with their images. We shall use this notion later in experimental part of this work.

\begin{define}[$name()$]
	Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x-y$.
\end{define}

\section{ID Set}
\label{section-definitions-id-set}

Based on the requirements for an \texttt{ID} attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.

\begin{define}[Candidate attribute mapping]
An attribute mapping $m$ is a~\textit{candidate attribute mapping} if it is an injective function, that is,
\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|.\]
\end{define}

\paragraph{Example}
In our example all the attribute mappings are candidate AMs.

Now we can proceed to define an ID set.

\begin{define}[ID set]
A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset.\]
\end{define}

That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one \texttt{ID} attribute).

\paragraph{Example}
Returning to our example, the following are all the possible ID sets: $\{ M_{y}^{a} \}, \{ M_{y}^{b} \}, \{ M_{y}^{b}, M_{z}^{a} \}, \{ M_{y}^{c} \}, \{ M_{y}^{c}, M_{z}^{a} \}$. Note that once we select an AM of type $y$ we can never add any other with the same type. Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.

\subsection*{\texttt{IDREF} and \texttt{IDREFS} Condition}

Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}:
\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i).\]
Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.

\section{Attribute Mapping Weight}
\label{section-definitions-weight}

This definition of weight for AMs or AM sets comes from \cite{fidax} again. Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.

\subsection{Support}

\begin{define}[Support]
\textit{Support} of an attribute mapping $m$ is defined as follows:
\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
\end{define}

\begin{quote}
The support of attribute mapping $M_x^y$ is the fraction of edges in the XML tree that connect $x$ elements to $y$ attributes.
\end{quote}

\paragraph{Example}
Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.\\

\subsection{Coverage}

\begin{define}[Coverage]
\textit{Coverage} of an attribute mapping $m$ is defined as follows:
\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
\end{define}

\begin{quote}
The coverage of an attribute mapping measures how much of the image of that mapping occurs elsewhere, as a fraction of all mappings images in the document.
\end{quote}

\paragraph{Example}
Coverage of $M_{y}^{a}$ in our example is $(0+0+1) / (2+1+1+1) = 0.2$. Coverage of $M_{z}^{a}$ is $0.2$ as well, all the other mappings have coverage $0$.
\\

\textit{Weight} of an attribute mapping is then defined as a linear combination of its support and coverage.

\begin{define}[Weight]
For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows:
\[weight(m) = \alpha . \phi(m) + \beta . \chi(m).\]
For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs:
\[weight(S) = \sum_{m \in S} weight(m).\]
\end{define}

Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even for example defined interactively by the user.

\section{Independent Set}
\label{section-definitions-is}

\nomenclature{IS}{Independent Set}

We shall need the notion of an indepentent set (IS) of vertices in a graph and its weighted variant.

\begin{define}[Independent set]
Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff
\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E.\]
\end{define}

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied:
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
\end{define}

It is well known that finding the maximum weighted IS is an NP-hard optimization problem, \cite{JM1986425}.

\section{Linear Programming}

The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints. The formulation is usually called a \textit{linear program}.

It can be written in the following form:

\begin{eqnarray*}
\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
\mathbf{x} & \geqslant & 0 \\
\end{eqnarray*}
\[
\mathbf{x} =
\begin{bmatrix}
x_1 \\
x_2 \\
\ldots \\
x_n \\
\end{bmatrix},
\mathbf{c} = 
\begin{bmatrix}
c_1 \\
c_2 \\
\ldots \\
c_n \\
\end{bmatrix},
\mathbf{b} =
\begin{bmatrix}
b_1 \\
b_2 \\
\ldots \\
b_m \\
\end{bmatrix},
\]

where a minimization version is possible, too.

Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function. $\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$. Furthermore, \[\max_{x} f(x)\] means maximize the value of $f(x)$ by changing the value of $x$.\\

Another way to write this formulation is this:

\begin{eqnarray*}
\max_{x} z = \sum_{i=1}^{n} c_i x_i \\
s.t. & \\
	   a_{11}x_{1} + a_{12}x_{2} + \ldots + a_{1n}x_{n} & \leqslant & b_{1} \\
	   a_{21}x_{1} + a_{22}x_{2} + \ldots + a_{2n}x_{n} & \leqslant & b_{2} \\
	   \ldots \\
	   a_{m1}x_{1} + a_{m2}x_{2} + \ldots + a_{mn}x_{n} & \leqslant & b_{m} \\
	   x_i \geqslant 0, i = 1, \ldots n \\
\end{eqnarray*}

Solving a linear program is usually possible in polynomial time using the \textit{simplex algorithm} described for example in \cite{dantzig1998linear}.

\section{Mixed Integer Problem}

\nomenclature{MIP}{Mixed Integer Problem}

\begin{define}[Mixed integer problem]
	MIP, or \textit{mixed integer problem}, is an~instance of~linear programming in~which some or~all variables are limited to~integral or~boolean (0, 1) values.
\end{define}

Solving MIP in general is $\NP$-hard.