### 2012-03-22 doc/theses/mikula/appendix_b.tex
Master thesis: In progress.

\chapter{Test Data} \label{APPENDIX_test_data}

Test data from the XMark project \cite{xmark}. Using the provided XML generator, XML document of size approximately 1.5 MB was generated. Figure \ref{listing_dtd} is its DTD and other figures in this appendix are XQuery queries that query the XML document.

The queries was slightly modified by replacing calls of \texttt{doc} function in paths by \texttt{/} (document node).

\begin{lstlisting}[caption=DTD of the test XML data, frame=single, label=listing_dtd]
<!-- DTD for auction database -->
<!-- $Id: auction.dtd,v 1.15 2001/01/29 21:42:35 albrecht Exp $ -->

<!ELEMENT site            (regions, categories, catgraph, people, open_auctions, closed_auctions)>

<!ELEMENT categories      (category+)>
<!ELEMENT category        (name, description)>
<!ATTLIST category        id ID #REQUIRED>
<!ELEMENT name            (#PCDATA)>
<!ELEMENT description     (text | parlist)>
<!ELEMENT text            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT bold            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT keyword         (#PCDATA | bold | keyword | emph)*>
<!ELEMENT emph            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT parlist         (listitem)*>
<!ELEMENT listitem        (text | parlist)*>

<!ELEMENT catgraph        (edge*)>
<!ELEMENT edge            EMPTY>
<!ATTLIST edge            from IDREF #REQUIRED to IDREF #REQUIRED>

<!ELEMENT regions         (africa, asia, australia, europe, namerica, samerica)>
<!ELEMENT africa          (item*)>
<!ELEMENT asia            (item*)>
<!ELEMENT australia       (item*)>
<!ELEMENT namerica        (item*)>
<!ELEMENT samerica        (item*)>
<!ELEMENT europe          (item*)>
<!ELEMENT item            (location, quantity, name, payment, description, shipping, incategory+, mailbox)>
<!ATTLIST item            id ID #REQUIRED
                          featured CDATA #IMPLIED>
<!ELEMENT location        (#PCDATA)>
<!ELEMENT quantity        (#PCDATA)>
<!ELEMENT payment         (#PCDATA)>
<!ELEMENT shipping        (#PCDATA)>
<!ELEMENT reserve         (#PCDATA)>
<!ELEMENT incategory      EMPTY>
<!ATTLIST incategory      category IDREF #REQUIRED>
<!ELEMENT mailbox         (mail*)>
<!ELEMENT mail            (from, to, date, text)>
<!ELEMENT from            (#PCDATA)>
<!ELEMENT to              (#PCDATA)>
<!ELEMENT date            (#PCDATA)>
<!ELEMENT itemref         EMPTY>
<!ATTLIST itemref         item IDREF #REQUIRED>
<!ELEMENT personref       EMPTY>
<!ATTLIST personref       person IDREF #REQUIRED>

<!ELEMENT people          (person*)>
<!ELEMENT person          (name, emailaddress, phone?, address?, homepage?, creditcard?, profile?, watches?)>
<!ATTLIST person          id ID #REQUIRED>
<!ELEMENT emailaddress    (#PCDATA)>
<!ELEMENT phone           (#PCDATA)>
<!ELEMENT address         (street, city, country, province?, zipcode)>
<!ELEMENT street          (#PCDATA)>
<!ELEMENT city            (#PCDATA)>
<!ELEMENT province        (#PCDATA)>
<!ELEMENT zipcode         (#PCDATA)>
<!ELEMENT country         (#PCDATA)>
<!ELEMENT homepage        (#PCDATA)>
<!ELEMENT creditcard      (#PCDATA)>
<!ELEMENT profile         (interest*, education?, gender?, business, age?)>
<!ATTLIST profile         income CDATA #IMPLIED>
<!ELEMENT interest        EMPTY>
<!ATTLIST interest        category IDREF #REQUIRED>
<!ELEMENT education       (#PCDATA)>
<!ELEMENT income          (#PCDATA)>
<!ELEMENT gender          (#PCDATA)>
<!ELEMENT business        (#PCDATA)>
<!ELEMENT age             (#PCDATA)>
<!ELEMENT watches         (watch*)>
<!ELEMENT watch           EMPTY>
<!ATTLIST watch           open_auction IDREF #REQUIRED>

<!ELEMENT open_auctions   (open_auction*)>
<!ELEMENT open_auction    (initial, reserve?, bidder*, current, privacy?, itemref, seller, annotation, quantity, type, interval)>
<!ATTLIST open_auction    id ID #REQUIRED>
<!ELEMENT privacy         (#PCDATA)>
<!ELEMENT initial         (#PCDATA)>
<!ELEMENT bidder          (date, time, personref, increase)>
<!ELEMENT seller          EMPTY>
<!ATTLIST seller          person IDREF #REQUIRED>
<!ELEMENT current         (#PCDATA)>
<!ELEMENT increase        (#PCDATA)>
<!ELEMENT type            (#PCDATA)>
<!ELEMENT interval        (start, end)>
<!ELEMENT start           (#PCDATA)>
<!ELEMENT end             (#PCDATA)>
<!ELEMENT time            (#PCDATA)>
<!ELEMENT status          (#PCDATA)>
<!ELEMENT amount          (#PCDATA)>

<!ELEMENT closed_auctions (closed_auction*)>
<!ELEMENT closed_auction  (seller, buyer, itemref, price, date, quantity, type, annotation?)>
<!ELEMENT buyer           EMPTY>
<!ATTLIST buyer           person IDREF #REQUIRED>
<!ELEMENT price           (#PCDATA)>
<!ELEMENT annotation      (author, description?, happiness)>

<!ELEMENT author          EMPTY>
<!ATTLIST author          person IDREF #REQUIRED>
<!ELEMENT happiness       (#PCDATA)>
\end{lstlisting}

\begin{lstlisting}[caption=Test query 1., frame=single, label=listing_test_query_1]
for $b in /site/people/person[@id = "person0"] return $b/name/text()
\end{lstlisting}

\begin{lstlisting}[caption=Test query 2., frame=single, label=listing_test_query_2]
for $b in /site/open_auctions/open_auction
return <increase>{$b/bidder[1]/increase/text()}</increase>
\end{lstlisting}

\begin{lstlisting}[caption=Test query 3., frame=single, label=listing_test_query_3]
for $b in /site/open_auctions/open_auction
where zero-or-one($b/bidder[1]/increase/text()) * 2 <= $b/bidder[last()]/increase/text()
return
  <increase
  first="{$b/bidder[1]/increase/text()}"
  last="{$b/bidder[last()]/increase/text()}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 4., frame=single, label=listing_test_query_4]
for $b in /site/open_auctions/open_auction
where
  some $pr1 in $b/bidder/personref[@person = "person20"],
       $pr2 in $b/bidder/personref[@person = "person51"]
  satisfies $pr1 << $pr2
return <history>{$b/reserve/text()}</history>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 5., frame=single, label=listing_test_query_5]
count(
  for $i in /site/closed_auctions/closed_auction
  where $i/price/text() >= 40
  return $i/price
)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 6., frame=single, label=listing_test_query_6]
for $b in //site/regions return count($b//item)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 7., frame=single, label=listing_test_query_7]
for $p in /site
return
  count($p//description) + count($p//annotation) + count($p//emailaddress)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 8., frame=single, label=listing_test_query_8]
for $p in /site/people/person
let $a :=
  for $t in /site/closed_auctions/closed_auction
  where $t/buyer/@person = $p/@id
  return $t
return <item person="{$p/name/text()}">{count($a)}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 9., frame=single, label=listing_test_query_9]
let $ca := /site/closed_auctions/closed_auction return
let
    $ei := /site/regions/europe/item
for $p in /site/people/person
let $a :=
  for $t in $ca
  where $p/@id = $t/buyer/@person
  return
    let $n := for $t2 in $ei where $t/itemref/@item = $t2/@id return $t2
    return <item>{$n/name/text()}</item>
return <person name="{$p/name/text()}">{$a}</person>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 10., frame=single, label=listing_test_query_10]
for $i in
  distinct-values(/site/people/person/profile/interest/@category)
let $p :=
  for $t in /site/people/person
  where $t/profile/interest/@category = $i
  return
    <personne>
      <statistiques>
        <sexe>{$t/profile/gender/text()}</sexe>
        <age>{$t/profile/age/text()}</age>
        <education>{$t/profile/education/text()}</education>
        <revenu>{fn:data($t/profile/@income)}</revenu>
      </statistiques>
      <coordonnees>
        <nom>{$t/name/text()}</nom>
        <rue>{$t/address/street/text()}</rue>
        <ville>{$t/address/city/text()}</ville>
        <pays>{$t/address/country/text()}</pays>
        <reseau>
          <courrier>{$t/emailaddress/text()}</courrier>
          <pagePerso>{$t/homepage/text()}</pagePerso>
        </reseau>
      </coordonnees>
      <cartePaiement>{$t/creditcard/text()}</cartePaiement>
    </personne>
return <categorie>{<id>{$i}</id>, $p}</categorie>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 11., frame=single, label=listing_test_query_11]
for $p in /site/people/person
let $l :=
  for $i in /site/open_auctions/open_auction/initial
  where $p/profile/@income > 5000 * exactly-one($i/text())
  return $i
return <items name="{$p/name/text()}">{count($l)}</items>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 12., frame=single, label=listing_test_query_12]
for $p in /site/people/person
let $l :=
  for $i in /site/open_auctions/open_auction/initial
  where $p/profile/@income > 5000 * exactly-one($i/text())
  return $i
where $p/profile/@income > 50000
return <items person="{$p/profile/@income}">{count($l)}</items>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 13., frame=single, label=listing_test_query_13]
for $i in /site/regions/australia/item
return <item name="{$i/name/text()}">{$i/description}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Testing query 14., frame=single, label=listing_test_query_14]
for $i in /site//item
where contains(string(exactly-one($i/description)), "gold")
return $i/name/text()
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 15., frame=single, label=listing_test_query_15]
for $a in
  /site/closed_auctions/closed_auction/annotation/description/parlist/
   listitem/
   parlist/
   listitem/
   text/
   emph/
   keyword/
   text()
return <text>{$a}</text>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 16., frame=single, label=listing_test_query_16]
for $a in /site/closed_auctions/closed_auction
where
  not(
    empty(
      $a/annotation/description/parlist/listitem/parlist/listitem/text/emph/
       keyword/
       text()
    )
  )
return <person id="{$a/seller/@person}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 17., frame=single, label=listing_test_query_17]
for $p in /site/people/person
where empty($p/homepage/text())
return <person name="{$p/name/text()}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 18., frame=single, label=listing_test_query_18]
declare namespace local = "http://www.foobar.org";
declare function local:convert($v as xs:decimal?) as xs:decimal?
{
  2.20371 * $v (: convert Dfl to Euro :)
};

for $i in /site/open_auctions/open_auction
return local:convert(zero-or-one($i/reserve))
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 19., frame=single, label=listing_test_query_19]
for $b in /site/regions//item
let $k := $b/name/text()
order by zero-or-one($b/location) ascending empty greatest
return <item name="{$k}">{$b/location/text()}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 20., frame=single, label=listing_test_query_20]
<result>
  <preferred>
    {count(/site/people/person/profile[@income >= 100000])}
  </preferred>
  <standard>
    {
      count(
        /site/people/person/
         profile[@income < 100000 and @income >= 30000]
      )
    }
  </standard>
  <challenge>
    {count(/site/people/person/profile[@income < 30000])}
  </challenge>
  <na>
    {
      count(
        for $p in /site/people/person
        where empty($p/profile/@income)
        return $p
      )
    }
  </na>
</result>
\end{lstlisting}