### 2012-03-31 doc/theses/mikula/appendix_a.tex
Master thesis: In progress.

\chapter{Attachments}
\begin{figure}[float=htpb]
Logical: \texttt{AND}, \texttt{OR}.

Comparison: \texttt{GEN\_EQUALS}, \texttt{GEN\_NOT\_EQUALS}, \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{VAL\_EQUALS}, \texttt{VAL\_NOT\_EQUALS}, \texttt{VAL\_LESS\_THAN}, \texttt{VAL\_LESS\_THAN\_EQUALS}, \texttt{VAL\_GREATER\_THAN}, \texttt{VAL\_GREATER\_THAN\_EQUALS}, \texttt{NOD\_IS}, \texttt{NOD\_PRECEDES}, \texttt{NOD\_FOLLOWS}.

Range: \texttt{TO}.

Additive: \texttt{PLUS}, \texttt{MINUS}.

Multiplicative: \texttt{MUL}, \texttt{DIV}, \texttt{IDIV}, \texttt{MOD}.

Set: \texttt{UNION}, \texttt{INTERSECTION}, \texttt{DIFFERENCE}.

Type test: \texttt{INSTANCE\_OF}, \texttt{CASTABLE\_AS}.

Type conversion: \texttt{TREAT\_AS}, \texttt{CAST\_AS}.

Unary: \texttt{UNARY\_PLUS}, \texttt{UNARY\_MINUS}.
\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
\label{FIG_operators}
\end{figure}

\begin{lstlisting}[caption=Resulting XSD of Test Scenario B, frame=single, label=listing_xsd]
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">
<!-- Inferred on Sat Mar 31 21:28:32 CEST 2012 by Basic IG generator, TwoStep(Iname (with attributes), Automaton Merging State(GreedyMDL(Combined(k,h-context, s,k-strings, Null, Null),Naive Alphabet), State Removal Ordered(Weighted)), Chained(Empty Children, Nested Concatenation, Null)), Basic XSD exporter -->

<!-- global types -->
<xs:complexType name="Temph" mixed="true">
  <xs:sequence>
    <xs:sequence minOccurs="0" maxOccurs="unbounded">
      <xs:choice>
        <xs:element name="keyword" type="Tkeyword"/>
        <xs:element name="bold" type="Tbold"/>
      </xs:choice>
    </xs:sequence>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tbold" mixed="true">
  <xs:sequence>
    <xs:choice>
      <xs:choice>
        <xs:sequence>
          <xs:element name="emph" type="Temph"/>
        </xs:sequence>
      </xs:choice>
      <xs:sequence>
        <xs:element name="keyword" type="Tkeyword"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tkeyword" mixed="true">
  <xs:sequence>
    <xs:choice>
      <xs:sequence minOccurs="0">
        <xs:element name="bold" type="Tbold"/>
      </xs:sequence>
      <xs:sequence>
        <xs:element name="emph" type="Temph"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Ttext" mixed="true">
  <xs:choice minOccurs="0" maxOccurs="unbounded">
    <xs:element name="bold" type="Tbold"/>
    <xs:element name="keyword" type="Tkeyword"/>
    <xs:element name="emph" type="Temph"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tparlist">
  <xs:sequence>
    <xs:element name="listitem" type="Tlistitem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tlistitem">
  <xs:choice>
    <xs:element name="text" type="Ttext"/>
    <xs:element name="parlist" type="Tparlist"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tdescription">
  <xs:choice>
    <xs:element name="parlist" type="Tparlist"/>
    <xs:element name="text" type="Ttext"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tincategory">
  <xs:attribute name="category" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tmail">
  <xs:sequence>
    <xs:element name="from" type="xs:string"/>
    <xs:element name="to" type="xs:string"/>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="text" type="Ttext"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tmailbox">
  <xs:sequence minOccurs="0">
    <xs:element name="mail" type="Tmail"/>
    <xs:element name="mail" type="Tmail" minOccurs="0"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Titem">
  <xs:sequence>
    <xs:element name="location" type="xs:string"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="payment" type="xs:string"/>
    <xs:element name="description" type="Tdescription"/>
    <xs:element name="shipping" type="xs:string"/>
    <xs:element name="incategory" type="Tincategory" minOccurs="0" maxOccurs="unbounded"/>
    <xs:element name="mailbox" type="Tmailbox"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tafrica">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tasia">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Taustralia">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Teurope">
  <xs:sequence>
    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tnamerica">
  <xs:sequence>
    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tsamerica">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tregions">
  <xs:sequence>
    <xs:element name="africa" type="Tafrica"/>
    <xs:element name="asia" type="Tasia"/>
    <xs:element name="australia" type="Taustralia"/>
    <xs:element name="europe" type="Teurope"/>
    <xs:element name="namerica" type="Tnamerica"/>
    <xs:element name="samerica" type="Tsamerica"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tcategory">
  <xs:sequence>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="description" type="Tdescription"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tcategories">
  <xs:sequence>
    <xs:element name="category" type="Tcategory"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tedge">
  <xs:attribute name="from" type="xs:string" use="required"/>
  <xs:attribute name="to" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tcatgraph">
  <xs:sequence>
    <xs:element name="edge" type="Tedge"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Taddress">
  <xs:sequence>
    <xs:element name="street" type="xs:string"/>
    <xs:element name="city" type="xs:string"/>
    <xs:element name="country" type="xs:string"/>
    <xs:choice>
      <xs:element name="zipcode" type="xs:string"/>
      <xs:sequence>
        <xs:element name="province" type="xs:string"/>
        <xs:element name="zipcode" type="xs:string"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tinterest">
  <xs:attribute name="category" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tprofile">
  <xs:sequence>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="education" type="xs:string"/>
      <xs:element name="interest" type="Tinterest"/>
    </xs:choice>
    <xs:choice>
      <xs:element name="business" type="xs:string"/>
      <xs:sequence>
        <xs:element name="gender" type="xs:string"/>
        <xs:element name="business" type="xs:string"/>
      </xs:sequence>
    </xs:choice>
    <xs:element name="age" type="xs:string" minOccurs="0"/>
  </xs:sequence>
  <xs:attribute name="income" type="xs:integer" use="required"/>
</xs:complexType>

<xs:complexType name="Twatch">
  <xs:attribute name="open_auction" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Twatches">
  <xs:sequence>
    <xs:element name="watch" type="Twatch" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tperson">
  <xs:sequence>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="emailaddress" type="xs:string"/>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="address" type="Taddress"/>
      <xs:element name="phone" type="xs:string"/>
      <xs:element name="homepage" type="xs:string"/>
      <xs:element name="creditcard" type="xs:string"/>
    </xs:choice>
    <xs:choice>
      <xs:sequence minOccurs="0">
        <xs:element name="profile" type="Tprofile"/>
        <xs:element name="watches" type="Twatches" minOccurs="0"/>
      </xs:sequence>
      <xs:element name="watches" type="Twatches"/>
    </xs:choice>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tpeople">
  <xs:sequence>
    <xs:element name="person" type="Tperson" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tpersonref">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tbidder">
  <xs:sequence>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="time" type="xs:string"/>
    <xs:element name="personref" type="Tpersonref"/>
    <xs:element name="increase" type="xs:integer"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Titemref">
  <xs:attribute name="item" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tseller">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tauthor">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tannotation">
  <xs:sequence>
    <xs:element name="author" type="Tauthor"/>
    <xs:element name="description" type="Tdescription"/>
    <xs:element name="happiness" type="xs:string"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tinterval">
  <xs:sequence>
    <xs:element name="start" type="xs:string"/>
    <xs:element name="end" type="xs:string"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Topen_auction">
  <xs:sequence>
    <xs:element name="initial" type="xs:integer"/>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="reserve" type="xs:decimal"/>
      <xs:element name="bidder" type="Tbidder"/>
    </xs:choice>
    <xs:element name="current" type="xs:string"/>
    <xs:element name="privacy" type="xs:string" minOccurs="0" maxOccurs="unbounded"/>
    <xs:element name="itemref" type="Titemref"/>
    <xs:element name="seller" type="Tseller"/>
    <xs:element name="annotation" type="Tannotation"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="type" type="xs:string"/>
    <xs:element name="interval" type="Tinterval"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Topen_auctions">
  <xs:sequence>
    <xs:element name="open_auction" type="Topen_auction" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tbuyer">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tclosed_auction">
  <xs:sequence>
    <xs:element name="seller" type="Tseller"/>
    <xs:element name="buyer" type="Tbuyer"/>
    <xs:element name="itemref" type="Titemref"/>
    <xs:element name="price" type="xs:integer"/>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="type" type="xs:string"/>
    <xs:element name="annotation" type="Tannotation"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tclosed_auctions">
  <xs:sequence>
    <xs:element name="closed_auction" type="Tclosed_auction" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tsite">
  <xs:sequence>
    <xs:element name="regions" type="Tregions"/>
    <xs:element name="categories" type="Tcategories"/>
    <xs:element name="catgraph" type="Tcatgraph"/>
    <xs:element name="people" type="Tpeople"/>
    <xs:element name="open_auctions" type="Topen_auctions"/>
    <xs:element name="closed_auctions" type="Tclosed_auctions"/>
  </xs:sequence>
</xs:complexType>

<!-- top level element -->
<xs:element name="site" type="Tsite">
  <xs:key name="key1">
    <xs:selector xpath="people/person"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:key name="key2">
    <xs:selector xpath=".//item"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:key name="key3">
    <xs:selector xpath="regions/europe/item"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:keyref name="keyRef1_key1" refer="key1">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="buyer/@person"/>
  </xs:keyref>
  <xs:keyref name="keyRef2_key2" refer="key2">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
  <xs:keyref name="keyRef3_key3" refer="key3">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
  <xs:keyref name="keyRef4_key3" refer="key3">
    <xs:selector xpath="open_auctions/open_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
</xs:element>
</xs:schema>
\end{lstlisting}