### 2012-04-12 doc/theses/mikula/appendix_c.tex
Master thesis: Adding forgotten appendix.

\chapter{Content of CD}
The CD attached to this thesis has the following structure.

\begin{itemize}
\item \texttt{content.txt} - A file with this text.
\item \texttt{text/} - A PDF version of the thesis.
\item \texttt{src/} - Source codes of the jInfer framework including implementation of our method. The same source codes can be also obtained from public Subversion repository by issuing command:\\
\texttt{svn co -r 2155 \\ https://jinfer.svn.sourceforge.net/svnroot/jinfer/jinfer/trunk/}
\item \texttt{bin/} - jInfer plugins for NetBeans 7.0.1. See jInfer Tutorial \cite{jinfer_tutorial} for step-by-step instructions, but use NetBeans 7.0.1. Since the tutorial is for the last official jInfer release 1.0, it says the required version of NetBeans is at least 6.9. But this is not true in our case, because we use the development version of jInfer and it requires NetBeans 7.0.1.
\item \texttt{testing/} - A directory containing sets of test data and test results. Again, see the Jinfer Tutorial \cite{jinfer_tutorial} for instructions how to run the inference. 
\end{itemize}